### 2012-12-08 docs/paper589.tex
Initial skeleton of the paper.

\documentclass[a4paper,12pt]{article}
\usepackage{amsmath} % For 'cases' curly brace
\usepackage{amsfonts} % For bold number set font
\usepackage{color}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{amssymb}

\newcommand{\xxx}[1]{\textcolor{red}{#1}}

\begin{document}
\title{Validating Cell Tower Clustering for Mobility Prediction}
\author{Mehrdad Moradi\\moradi@umich.edu \and Pedro d'Aquino\\pdaquino@umich.edu}
\date{December 11, 2012}

\maketitle

\section{Introduction}

\xxx{Talk about mobility prediction in mobile networks. Discuss some of the potential benefits -- some of them were discussed in the
ATT leap graph paper. Maybe do a little research on this.}

\cite{Jadbabaie2012}
\cite{DeGroot1974}

\section{Motivation}

\xxx{This is where we explain why we think that our work is important: because current tower clustering methods do not have ground truth data to check against and we do.}

\section{Methodology}

\xxx{What were out methods?}

\section{Clustering algorithm}

\xxx{Explain why we used the Mobility Profiler algorithm; mention that we couldn't use a leap graph because we don't have active set information.
\\Describe the clustering algorithm.}

\subsection{GSM datasets}

\xxx{MDC and Reality Mining.}

\subsection{Cell location datasets}

\xxx{GPS logs from MDC and OpenCell}

\section{Results}

\subsection{Dataset comparison}

\xxx{Compare MDC and Reality Mining, i.e. show that they are similiar using the sampled RM?}

\subsection{Location comparison}

\bibliographystyle{plain}
\bibliography{eecs589Report}

\end{document}

### 2012-12-10 src/edu/umich/eecs/data/ClusterGenerator.java
Removed wrongly

@@ -20,7 +20,6 @@ import edu.umich.eecs.util.Tic;
 
 public class ClusterGenerator {
 	public static Tic clock = new Tic(true);
-	
 
 	public static void computeCluster(DataSetType dataset){
 		OscillationService os= new OscillationService(dataset);
@@ -45,8 +44,10 @@ public class ClusterGenerator {
 	
 	public static void main(String[] args) {
 		
+		ClusterGenerator.computeCluster(DataSetType.SampledRealityMining);
 		ClusterGenerator.computeCluster(DataSetType.RealityMining);
-
+		ClusterGenerator.computeCluster(DataSetType.NokiaChallenge);
+		
 		
 	}
 

### 2012-12-10 src/edu/umich/eecs/data/ClusterGenerator.java
resolve error

@@ -20,7 +20,6 @@ import edu.umich.eecs.util.Tic;
 
 public class ClusterGenerator {
 	public static Tic clock = new Tic(true);
-	
 
 	public static void computeCluster(DataSetType dataset){
 		OscillationService os= new OscillationService(dataset);
@@ -45,9 +44,9 @@ public class ClusterGenerator {
 	
 	public static void main(String[] args) {
 		
+		ClusterGenerator.computeCluster(DataSetType.SampledRealityMining);
 		ClusterGenerator.computeCluster(DataSetType.RealityMining);
-
-		
+		ClusterGenerator.computeCluster(DataSetType.NokiaChallenge);	
 	}
 
 }

### 2012-12-10 src/edu/umich/eecs/data/DensityGenerator.java

@@ -16,6 +16,8 @@ import edu.umich.eecs.service.CellService;
 import edu.umich.eecs.service.CellSpanService;
 import edu.umich.eecs.service.CellSpanServiceInterface;
 import edu.umich.eecs.service.ClusterService;
+import edu.umich.eecs.service.MDCCellSpanService;
+import edu.umich.eecs.service.SampledCellSpanService;
 
 public class DensityGenerator {
 	
@@ -61,7 +63,10 @@ public class DensityGenerator {
 	
 	public static void main(String[] args) {
 		
+		new DensityGenerator().computeDensity(new MDCCellSpanService(), DataSetType.NokiaChallenge);
 		new DensityGenerator().computeDensity(new CellSpanService(), DataSetType.RealityMining);
+		new DensityGenerator().computeDensity(new SampledCellSpanService(), DataSetType.SampledRealityMining);
+		
 	}
 
 }

### 2012-12-10 src/edu/umich/eecs/DumpClusterFinder.java
Adding Quality and Size Comparison Figs

@@ -1,0 +1,21 @@
+package edu.umich.eecs;
+
+import java.util.List;
+
+import edu.umich.eecs.dto.Cluster;
+import edu.umich.eecs.dto.DataSetType;
+import edu.umich.eecs.service.ClusterService;
+
+public class DumpClusterFinder {
+
+	
+	List< Cluster > clusters;       ///clusterset after running clustering algorithms.
+	ClusterService clusterSrv;
+	Integer transNumber;// includes number of package of 100 clusters that persisted
+	DataSetType dataSet; //dataset type the cluster is creating from.
+	
+	
+	
+	
+	
+}

### 2012-12-10 src/edu/umich/eecs/statistics/CellListParser.java

@@ -13,7 +13,7 @@ public class CellListParser {
 	
 	/**
 	 * Reads a list of cells that are specified like this:
-	 * CELLID   LAC   MNC   MCC
+	 * CELLID LAC MNC MCC
 	 * The method assumes there is no header.
 	 * @param scanner
 	 * @return

### 2012-12-10 src/edu/umich/eecs/statistics/CellLocationStatistics.java

@@ -62,10 +62,10 @@ public class CellLocationStatistics {
 		// query and I didn't want to do it in Hibernate.
 		//
 		Scanner scanner = new Scanner(new File(
-				"src/edu/umich/eecs/statistics/files/cells_opencell_and_gpslog_all.txt"));
+				"/Users/Mehrdad/Documents/workspace/589-Project/files/cells_opencell_and_gpslog_all.txt"));
 		try {
 
-			// CELLID   LAC   MNC   MCC  
+			// CELLID LAC MNC MCC
 			scanner.nextLine();
 			Set<CellKey> cells = new HashSet<CellKey>(CellListParser.fromFile(scanner));
 			CellService cellSvc = new CellService();

### 2012-12-10 src/edu/umich/eecs/statistics/ClusterIncompletenessStatistics.java

@@ -145,7 +145,7 @@ public class ClusterIncompletenessStatistics {
 				new CellService(),
 				new ClusterService(), dataset);
 		
-		Scanner scanner = new Scanner(new File("src/edu/umich/eecs/statistics/files/clustered_cells_with_gpslog.txt"));
+		Scanner scanner = new Scanner(new File("/Users/Mehrdad/Documents/workspace/589-Project/files/clustered_cells_with_gpslog.txt"));
 		
 		scanner.nextLine(); // get rid of the header
 		List<CellKey> cellsWithGpsData = CellListParser.fromFile(scanner);

### 2012-12-10 src/edu/umich/eecs/statistics/ClusterQualityStatistics.java

@@ -84,7 +84,7 @@ public class ClusterQualityStatistics {
 		DataSetType dataset = DataSetType.NokiaChallenge;
 		
 		ClusterQualityStatistics qualityStats = new ClusterQualityStatistics(new ClusterService());
-		Scanner scanner = new Scanner(new File("src/edu/umich/eecs/statistics/files/clustered_cells_with_gpslog.txt"));
+		Scanner scanner = new Scanner(new File("/Users/Mehrdad/Documents/workspace/589-Project/files/clustered_cells_with_gpslog.txt"));
 		scanner.nextLine(); // get rid of the header
 		List<CellKey> cellsWithGpsData = CellListParser.fromFile(scanner);
 		qualityStats.printClusterLocationInfo(cellsWithGpsData,  new CellService(), dataset);

### 2012-12-10 src/edu/umich/eecs/DumpClusterFinder.java
Added Density Distribution

@@ -5,6 +5,11 @@ import java.util.List;
 import edu.umich.eecs.dto.Cluster;
 import edu.umich.eecs.dto.DataSetType;
 import edu.umich.eecs.service.ClusterService;
+/**
+ * @author Mehrdad
+ * In this class, we cluster cells based on areaid and country id.
+ *
+ */
 
 public class DumpClusterFinder {
 
@@ -13,6 +18,25 @@ public class DumpClusterFinder {
 	ClusterService clusterSrv;
 	Integer transNumber;// includes number of package of 100 clusters that persisted
 	DataSetType dataSet; //dataset type the cluster is creating from.
+	
+	
+
+	public DumpClusterFinder(List<Cluster> clusters, ClusterService clusterSrv,
+			Integer transNumber, DataSetType dataSet) {
+		super();
+		this.clusters = clusters;
+		this.clusterSrv = clusterSrv;
+		this.transNumber = transNumber;
+		this.dataSet = dataSet;
+	}
+	
+	public void Compute(){
+		
+		
+		
+	}
+	
+	
 	
 	
 	

### 2012-12-10 src/edu/umich/eecs/data/DensityGenerator.java

@@ -41,10 +41,11 @@ public class DensityGenerator {
 			
 			double numberOfCluster= cc.get(i)+uc.get(i);
 			double numberOfCells=ac.get(i);
-			double density= numberOfCells/numberOfCluster;
+			double density= numberOfCluster/numberOfCells;
 			AreaDensity ad= new AreaDensity(i, (int)numberOfCluster, dataset, (int)numberOfCells,density );
 			addDensity(ad);
 		}	
+		aSrv.saveListToAD(listAreaDensity);
 		System.out.println("#"+(50*numberOfPersistence+listAreaDensity.size())+" Area Densities is Persisted");
 		listAreaDensity.clear();
 	}

### 2012-12-10 src/edu/umich/eecs/dto/AreaDensity.java

@@ -3,12 +3,17 @@ package edu.umich.eecs.dto;
 import java.io.Serializable;
 
 import javax.persistence.Entity;
+import javax.persistence.GeneratedValue;
+import javax.persistence.GenerationType;
 import javax.persistence.Id;
 
 @Entity
 public class AreaDensity implements Serializable {
 	
 	@Id
+	@GeneratedValue(strategy=GenerationType.AUTO)
+	int id;
+	
 	int areaCode;
 	int numberOfCluster;
 	DataSetType dataSet;
@@ -24,7 +29,7 @@ public class AreaDensity implements Serializable {
 	int numberOfCells;
 	
 	/*
-	 * density=(number of cells in a specific area/ number of cluster in an area)
+	 * density=(number of cluster in an area/number of cells in a specific area )
 	 */
 	double density;
 	

### 2012-12-10 src/edu/umich/eecs/service/AreaDensityService.java

@@ -2,9 +2,14 @@ package edu.umich.eecs.service;
 
 import java.util.List;
 
+import org.hibernate.Query;
 import org.hibernate.Session;
 
 import edu.umich.eecs.dto.AreaDensity;
+import edu.umich.eecs.dto.Cell;
+import edu.umich.eecs.dto.CellKey;
+import edu.umich.eecs.dto.DataSetType;
+import edu.umich.eecs.dto.OscillatingCellTowerPair;
 
 
 public class AreaDensityService  extends Service{
@@ -24,6 +29,15 @@ public class AreaDensityService  extends Service{
 			} 
 			commitTransaction(s);
 		}
+		
+		public List<Double> getAllDensity(DataSetType dataset) {
+			  Session s= fireTransaction();
+			  Query query=s.createQuery("select density from AreaDensity  where dataset =:dataset");
+			  query.setInteger("dataset", dataset.asInt());
+			   List <Double> alldensity=(List<Double>)query.list();
+			  return alldensity;	
+		}
+		
 		
 
 

### 2012-12-10 src/edu/umich/eecs/statistics/ClusterStatistics.java

@@ -3,8 +3,10 @@ package edu.umich.eecs.statistics;
 import java.util.List;
 
 import edu.umich.eecs.data.Constants;
+import edu.umich.eecs.dto.AreaDensity;
 import edu.umich.eecs.dto.Cluster;
 import edu.umich.eecs.dto.DataSetType;
+import edu.umich.eecs.service.AreaDensityService;
 import edu.umich.eecs.service.ClusterService;
 
 /**
@@ -65,7 +67,31 @@ public class ClusterStatistics {
 		
 	}
 	
-	public void clusterDensity(){
+	public void clusterDensity(DataSetType dataset){
+		String name="";
+		switch (dataset) {
+		case RealityMining:
+			name="Reality";
+			break;
+		case SampledRealityMining:
+			name="SampledReality";
+			break;
+			
+		case NokiaChallenge:
+			name="NokiaChallange";
+			break;
+	
+		}
+		AreaDensityService srv= new AreaDensityService();
+		
+		StatisticFile area= new StatisticFile(dataset.asInt()+".txt");
+		
+		for(Double ad: srv.getAllDensity(dataset)){
+			area.writeDouble(ad);
+			area.writeString("\n");
+		}
+		
+		area.tearDown();
 		//
 	}
 	
@@ -74,10 +100,19 @@ public class ClusterStatistics {
 	
 	public static void main(String[] args) {
 		ClusterStatistics css= new ClusterStatistics();
-		css.sizeQuality(DataSetType.NokiaChallenge);
+		/*css.sizeQuality(DataSetType.NokiaChallenge);
 		css.sizeQuality(DataSetType.RealityMining);
 		css.sizeQuality(DataSetType.SampledRealityMining);
 		System.out.println("Quality and Size Metrics are Generated For All Dataset");
+		*/
+		
+		
+		css.clusterDensity(DataSetType.NokiaChallenge);
+		css.clusterDensity(DataSetType.RealityMining);
+		css.clusterDensity(DataSetType.SampledRealityMining);
+		System.out.println("Area Density is Created!....");
+
+		
 	}
 	
 	

### 2012-12-11 src/edu/umich/eecs/statistics/CellLocationStatistics.java
Added Figs

@@ -50,6 +50,7 @@ public class CellLocationStatistics {
 			
 		 //System.out.println(/*i++ + "/" + cells.size() + " (" + */disparity); //+ ")");
 			System.out.println(key.getCountryID() + "\t" + key.getNetworkID() + "\t" + key.getAreaID() + "\t" + key.getCellID() + "\t" + disparity);
+			System.out.println(disparity);
 		}
 		
 		return MeanAndStdDev.fromSample(disparities);
@@ -62,11 +63,12 @@ public class CellLocationStatistics {
 		// query and I didn't want to do it in Hibernate.
 		//
 		Scanner scanner = new Scanner(new File(
-				"/Users/Mehrdad/Documents/workspace/589-Project/files/cells_opencell_and_gpslog_all.txt"));
+				"/Users/Mehrdad/Documents/workspace/589-Project/files/cells_opencell_and_gpslog.txt"));
 		try {
 
 			// CELLID LAC MNC MCC
-			scanner.nextLine();
+			String s = scanner.nextLine();
+			System.out.println(s);
 			Set<CellKey> cells = new HashSet<CellKey>(CellListParser.fromFile(scanner));
 			CellService cellSvc = new CellService();
 			OpenCellService openCellSvc = new OpenCellService();

### 2012-12-11 src/edu/umich/eecs/statistics/ClusterIncompletenessStatistics.java

@@ -30,6 +30,7 @@ public class ClusterIncompletenessStatistics {
 	private ClusterService clusterSvc;
 	private Tic clock = new Tic(true);
 	private DataSetType dataset;
+	private Map<Integer, Integer> clusterToCountUnclutered= new HashMap<>();
 	private final double radius = 500;  // cells within radius meters of each other should be clustered.
 	
 	
@@ -47,6 +48,8 @@ public class ClusterIncompletenessStatistics {
 		int countTotalUnclusteredCells = 0;
 		int countValidCells = 0;
 		System.out.println("CountryID\tNetworkID\tAreaID\tCellID\tClusterID\tCountUnclustered");
+		//System.out.println("CellID\t\tClusterID\t\tCountUnclustered");
+
 		for(CellKey cell1 : cells) {
 			int countCellUnclusteredCells = 0;
 			if(!isValidCell(cell1, cellGpsMap)) {
@@ -85,6 +88,13 @@ public class ClusterIncompletenessStatistics {
 			+ cell1.getCellID() + "\t"
 			+ (cellClusterMap.containsKey(cell1) ? cellClusterMap.get(cell1).getCkey().getClusterID() : "-1")
 			+ "\t" + countCellUnclusteredCells);
+			
+			//System.out.println(countCellUnclusteredCells);
+			
+			
+			
+			
+			
 		}
 		double countUnclusteredCells = countTotalUnclusteredCells / 2; // our method double counts (c1, c2) and (c2, c1)
 		double averageCountUnclusteredCells = countUnclusteredCells / countValidCells;

### 2012-12-11 src/edu/umich/eecs/statistics/ClusterQualityStatistics.java

@@ -60,11 +60,17 @@ public class ClusterQualityStatistics {
 				}
 				MeanAndStdDev clusterDistances = MeanAndStdDev
 						.fromSample(distances);
-				System.out.println(cluster.getCkey().getClusterID() + "\t"
+				/*System.out.println(cluster.getCkey().getClusterID() + "\t"
 						+ clusterCells.size() + "\t"
 						+ gpsPositionsInCluster.size() + "\t"
 						+ clusterDistances.getMean() + "\t"
-						+ clusterDistances.getStdDev());
+						+ clusterDistances.getStdDev());*/
+				
+				/*double clusterSize=clusterCells.size();
+				double gpsPositionInCluster=gpsPositionsInCluster.size();*/
+				
+				System.out.println(clusterDistances.getStdDev());
+						
 			}
 		}
 	}

### 2012-12-11 docs/paper589.tex
Latest version of the paper.

@@
 \usepackage{amssymb}
+\usepackage{url}
 
@@
 
-\cite{Jadbabaie2012}
-\cite{DeGroot1974}
+In this project, we address one of the challenges faced in mobility prediction using GSM cell tower connection logs. In our context, mobility prediction in cellular
+towers can be defined as the ability to predict a user's physical movement by examning the history of cell tower connections. This is very important
+for mobile network operators, as several network optimizations are possible if they can forecast with reasonable accuracy
+where a user will be in the future \cite{LeapGraph} \cite{Liu97anoptimal}. By being able to pre-allocate resources to towers that are most likely to
+be connected to by the user, network operators can improve latency and uptime. This is increasingly important as the amount of data in cellular networks
+increases: indeed, in 2010 the amount of data traffic surpassed voice traffic for the first time, and it is expected to double annually in the near future
+\cite{EricssonData}.
+
+Another use of mobility prediction is to discover mobility profiles, i.e.
+patterns in a user's movement \cite{mobilityprofiler}.
+As illustrated in \cite{mobilityprofiler}, mobility profiles have a wide range of possible applications, from traffic planning to advertisement.
+Recently, Google started integrating
+a contextual search and recommendation service in its phones that is able to, for instance, detect a user's home and work addresses and suggest routes based on
+traffic \cite{googleNow}.
+
+The heart of every mobility prediction system based on cell tower connections is to infer movement based on which cell the user's phone is connected to. The goal
+is to predict which tower the user is more likely to connect to next.
+A major challenge in this is dealing with the \textit{oscillation effect}, which happens when a \textit{stationary} user's phone switches between two or more towers.
+This is a natural consequence of the radio medium and the fact that towers very often have overlapping ranges.
+In order to accurately predict a user's mobility, these outliers have to be removed so that fluctuations can be separated from actual user movement.
+
+In this project, we analyze one published algorithm for oscillation detection \cite{mobilityprofiler} and evaluate its performance using GPS data as ground-truth.
+
+\xxx{``In the next section, we describe....''}
+\xxx{More about the results.}
 
@@
 
+\xxx{How do we cite the Nokia MDC?}
+
+\xxx{Reproduce the stationary phone example in the leap graph paper.}
+
+The accuracy of current mobility prediction systems based on cell tower connections is usually measured through machine learning techniques like cross-validation that try to validate the ultimate goal of forecasting the user's future connections \cite{LeapGraph}.
+As far as we know there are no studies that evaluate methods of dealing with the oscillation effect, even though all prediction systems must deal with it.
+
+In this paper we report the results of such an analysis.
+We have implemented and evaluated the ``oscillation graph'' technique described in \cite{mobilityprofiler}.
+The basic idea of this algorithm is to detect cell towers that often present oscillations and consider them as a single tower.
+This is effectively a clustering algorithm.
+Our goal is to evaluate whether towers that are clustered together are actually close in real life.
+To evaluate this, we use the Nokia Mobility Challenge dataset, which has data on both cell tower connection (``GSM logs'') and the GPS position of the phones.
+
 \section{Methodology}
 
-\xxx{What were out methods?}
+\xxx{What were our methods?\\- Implemented clustering algorithm. Explain mobility paths, oscillations, oscillation graph etc. \\- RM. We sampled the RM dataset and compared it to
+the Nokia MDC \\- MDC. We apply the oscillation clustering to the MDC and look at the position
+characteristics of the clusters.}
 
-\section{Clustering algorithm}
+The Mobility Profiler paper \cite{mobilityprofiler} describes an algorithm to detect the oscillation effect.
+Our approach to evaluating this algorithm's efficiency was the following.
+We implement the algorithm as described in \cite{mobilityprofiler} and test it in the same dataset used in the original paper.
+However, that dataset does not have any data on the location of the users when they were connected to the towers.
+We therefore apply the algorithm to another dataset, from the Nokia Mobility Challenge, in which the GPS position of the users is listed.
+In order to evaluate how successful the clustering algorithm is, we devise two metrics: cluster distance and cluster incompleteness.
 
-\xxx{Explain why we used the Mobility Profiler algorithm; mention that we couldn't use a leap graph because we don't have active set information.
-\\Describe the clustering algorithm.}
+\subsection{Clustering algorithm}
+
+There are two main papers that tackle the problem of oscillation effects: the Mobility Profiler paper, by Bayir et al.
+\cite{mobilityprofiler}, and the Leap Graph paper by Duffield et al \cite{LeapGraph}.
+Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by resarchers at AT\&T, makes use of the active set of connections.
+The active set contains all towers that are within range of the phone's antenna at any given time.
+Unfortunately, active set logs are not public which makes it impossible for us to evaluate the leap graph algorithm.
+
+We therefore focus our efforts on the algorithm described in the Mobility Profiler paper, which we call the oscillation graph algorithm.
+In the interest of completeness, we briefly describe how it works.
+
+\xxx{Cellspan log example. Mobility path example.}
+
+\subsubsection{Mobility paths}
+The input of the algorithm is a sequence of \textit{handovers}, which are transition events when a phone switches the tower it is connected to.
+These events are timestamped.
+The first step of the algorithm is to analyze the transitions and extract \textit{mobility paths}.
+A mobility path corresponds to a sequence of handovers that are associated with actual user movement.
+For example, imagine a user that drives every morning to work, stays there for eight hours and goes back home.
+During his drive to work, the user's phone will switch cell towers several times.
+These handovers will be included in a mobility path.
+While the user is at work, ignoring the oscillation effect temporarily, the phone will be connected to a single tower.
+Finally, when the user the goes home the handovers that occur during the drive will belong to a different mobility path.
+
+More formally, let $\Delta_i$ be the time difference between two consecutive handovers for a particular user's connection log.
+In other words, $\Delta_i$ is the time the user spent connected to cell tower $C_i$.
+Let $\delta_{duration}$ be the \textit{end-location threshold}.
+Then we define a mobility path as a sequence of cell connections $\left\{C_i, C_{i+1}, \ldots, C_j\right\}$ such that $\Delta_i < \delta_{duration}$ for all handovers.\footnote{We also have to consider \textit{hidden end-locations}, which happen when a user's phone is disconnected from any towers, or if the user turns his phone off.
+We omit its formal definition for briefness.}
+
+\subsubsection{Oscillating pair detection}
+
+The goal of the oscillation detection algorithm is to retrieve sets of towers that overlap each other -- that is, towers that could be in a stationary user's handover logs.
+The main idea behind the algorithm is that a user's phone is more likely to switch between oscillating towers than between regular towers that are not overlapping.
+In this context, \textit{switching} means a handover from one tower to another inside of a mobility path.
+We can define $k$ to be the minimum switching count and consider all pairs of towers that have more than $k$ switches in one mobility path to be an \textit{oscillating pair}.
+The example in \cite{mobilityprofiler} is illustrative: consider a sequence of handovers $\left[x, y, x, w, v, w, y\right]$ where every element represents one cell tower.
+We say that the pair $(x, y)$ has 3 switches: the first twofrom $x$ to $y$ and back to $x$ in the first positions of the sequence, and then back to $y$ at the end.
+Hence, if we adopt $k=3$ the only oscillating pair would be $(x, y)$.
+Note that switches do not need to be consecutive.
+The authors argue that this allows for detection of overlapping towers in dense region where multiple towers may be within range at the same time.
+
+\subsubsection{Oscillation graph}
+
+After detecting all the oscillating pairs in all mobility paths, the next step is to build the oscillation graph.
+The set $V$ of vertices in the graph are all cell towers in the dataset, and the weighted edges in $E$ connect oscillating pairs.
+Let $P_i$ be a mobility path.The weight of every edge, which we also call its \textit{support}, is defined as:
+
+\begin{equation*}
+s(x,y) = \frac{|\left\{P_i| (x, y) \in P_i \wedge (x, y) \text{oscillated}\right\}|}{|\left\{P_i| (x, y) \in P_i\right\}|}
+\end{equation*}
+
+In words, $s(x,y)$ is the ratio of how many times the cell towers $x$ and $y$ oscillated over the number of mobility paths in which $x$ and $y$ appear.
+For example, a pair of towers such that in \textit{every} mobility path that both of them appear the switch count is greater than $k$ would have a support equal to 1.
+If a pair only switches more or equal to $k$ times in half the mobility paths both of the towers appear in, that pair's support would be 0.5.
+
+\subsubsection{Oscillation graph clustering}
+
+\xxx{Explain the clustering algorithm.}
+
+The final step of the algorithm is to find clusters in the oscillation graph.
+Intuitively, a cluster represents a group of towers that are not indicative of actual user movement.
+The Mobility Profiler paper uses an greedy divise algorithm for clustering.
+At every iteration, the algorithm removes the edge with the lowest weight from the graph.
+It then evaluates the ``quality'' of all connected components, which is the weighted ratio of edges inside of the cluster over the number of edges that leave the cluster in the original graph.
+Formally, let $C$ be a cluster with the set of edges $E$, and $E_{out}$ be the set of edges that have exactly one vertex in $C$, and define $w(e)$ to be an edge's weight:
+
+\begin{equation*}
+Q(C) = \frac{\displaystyle\sum_{\forall e_{in} \in E}{w(e_{in})}}{\displaystyle\sum_{\forall e_{out} \in E_{out}}{w(e_{out})}}
+\end{equation*}
+
+Every cluster with quality metric greater than a threshold is removed from the graph.
+The algorithm continues until there are no more vertices in the graph.
+The algorithm outputs all the clusters it found.
+Our main goal is to evaluate these clusters according to the GPS position of the users when they were within range of that tower.
 
@@
 
-\xxx{MDC and Reality Mining.}
+\xxx{Reality Mining dataset reference.}
+
+We use two datasets with GSM cell tower connection data, from the Reality Mining project and from the Nokia Mobility Data Challenge.
+
+\subsubsection{Reality Mining}
+
+\xxx{Describe the reality mining dataset}
+
+\subsubsection{Nokia Mobility Data Challenge -- GSM}
+
+\xxx{Describe the Nokia Mobility Challenge}
 
@@
 
+Our main contribution is an analysis of the oscillation clustering algorithm using GPS data. We use two sources of GPS data, the Nokia MDC and the OpenCell database.
+
+\subsubsection{Nokia Mobility Data Challenge -- GPS}
+
+Phones in the MDC also had their GPS units sampled every minute.
+The format of the GPS logs is shown in Figure \xxx{Show the MDC GPS logs}.
+We process the GPS and try to establish an approximate position for each cell tower.
+We approximate the cell tower position as the average of the positions of users when they were connected to that tower.
+To do this, we have to match each entry in the GSM log with an entry in the GPS log.
+
+Because GSM and GPS entries do not necessarily have the same timestamp (in fact, they almost never do), we use GPS entries that have a timestamp within $\delta_{GPS}$ seconds of the GSM entry we are trying to match.
+The algorithm goes through every record in the GSM log and tries to find a matching GPS log.
+After going through all users' connection logs, we estimate each tower's position as the average of all GPS positions that were associated with that tower.
+It is important to have a measure of how accurate the GPS estimate of a particular tower is, so we record the number of GPS samples for that tower and the standard deviation in the samples.
+The standard deviation is computed based on the distance from each sample to the average point, which is calculated using the Haversine formula \cite{haversine}.
+
+The algorithm runs in $O(k(n\log n + m\log n))$, where $k$ is the number of users in the dataset and $m$ and $n$ are upper bounds on the number of GPS and GSM records per user, respectively.
+The logarithmic factor comes from sorting the GPS log by timestamp and performing a binary search.
+
+\subsubsection{OpenCell}
+
+\xxx{Talk about OpenCell; what it is and how we use it.}
+
 \section{Results}
+
+\xxx{Results introduction fluff.}
+
+\xxx{Talk about challenges in handling the data -- wrong clock, entries out of order in the RM dataset;}
+
+We implemented the clustering algorithm in \cite{mobilityprofiler} for the Reality Mining and the Nokia Mobility Data Challenge datasets and compared the results of the clustering in both datasets (adjusting for the fact that the MDC dataset's data is sampled every minute).
+In this section, we report the results of running the algorithm in those datasets and analyze the results.
+\xxx{One sentence description of the non-GPS related results.} We also used GPS data to evaluate the accuracy and quality of the resulting clusters.
+\xxx{One sentence description of the GPS results}.
+
+Running the clustering algorithm on the RM dataset was more challenging than originally expected; there were inconsistencies with the dates (apparently the phones' clocks reverted to 01/01/2004 when the battery died) and time (with entries appearing in an inconsistent order; see \xxx{RM GSM time weirdness example}).
+We dealt with these problems by ignoring records with a timestamp equal to 01/01/2004 and by assuming that the ordering of the lines in the dataset logs represented the actual ordering that the logs were collected -- this seemed consistent with manual analysis of the logs and generated valid mobility paths, which is not the case if the timestamps are taken at face value.
+\xxx{Maybe expand on what a valid mobility path is.}
 
@@
 
+\section{Conclusion}
+
+\xxx{Talk about other clustering techinques and other modularity measures, like Newman's.}
+
 \bibliographystyle{plain}

### 2012-12-11 src/edu/umich/eecs/data/CellGpsDataImporter.java
Merge.

@@ -238,8 +238,8 @@ public class CellGpsDataImporter {
 		// Replace this with the path to where the personid/gsm.csv files are.
 		//
 		
-		String pathToCsvFiles ="/Users/Mehrdad/Documents/workspace/589-Project/mdc_csv/mdc2012-373-taskopen/";
-		//String pathToCsvFiles = "C:\\Users\\Pedro\\Desktop\\589 project data\\mdc2012-373-taskopen\\mdc2012-373-taskopen\\";
+		//String pathToCsvFiles ="/Users/Mehrdad/Documents/workspace/589-Project/mdc_csv/mdc2012-373-taskopen/";
+		String pathToCsvFiles = "C:\\Users\\Pedro\\Desktop\\589 project data\\mdc2012-373-taskopen\\mdc2012-373-taskopen\\";
 		//
 		// Every person has a directory with her own ID. Inside that directory are files called
 		// "gsm.csv" and "gps.csv" with the data we want.

### 2012-12-11 src/edu/umich/eecs/statistics/CellLocationStatistics.java

@@ -63,7 +63,8 @@ public class CellLocationStatistics {
 		// query and I didn't want to do it in Hibernate.
 		//
 		Scanner scanner = new Scanner(new File(
-				"/Users/Mehrdad/Documents/workspace/589-Project/files/cells_opencell_and_gpslog.txt"));
+				//"/Users/Mehrdad/Documents/workspace/589-Project/files/cells_opencell_and_gpslog.txt"));
+				"src/edu/umich/eecs/statistics/files/cells_opencell_and_gpslog.txt"));
 		try {
 
 			// CELLID LAC MNC MCC

### 2012-12-12 docs/paper589.tex
More info.

@@
-\documentclass[a4paper,12pt]{article}
+\documentclass[a4paper,10pt]{article}
 \usepackage{amsmath} % For 'cases' curly brace
@@
 
+In all experiments we used the same parameters value as the original Mobility Profiler paper, namely \xxx{list the constants we used, like the minimum oscillation count etc.}.
+
 \subsection{Dataset comparison}
@@
 
-\subsection{Location comparison}
+We compare the Reality Mining dataset, which was originally used in the Mobility Profiler paper, with the Mobility Data Challenge.
+Table \xxx{Table with number of cells etc.} shows the number of cells, area codes and countries that are represented in each dataset.
+We can see that the MDC data has a much larger breadth in terms of the countries that are covered, but contains less cells and cell transitions.
+\xxx{Try to explain this difference.
+Talk about number of clusters.}
+
+However, both datasets show similar patterns in cluster size distribution, as evidenced in Figure \xxx{Figure with the oscillation histogram}, which also presents the cluster size distribution for the Reality Mining dataset sampled every minute.
+This is necessary for a fair comparison since the RM contains information on every handover, while the MDC samples the tower the phone is currently connected to every minute (so there is loss of information).
+
+The fact that both datasets show a similar distribution is important for our results, since it allows us to extrapolate the results we find with the MDC dataset to the RM dataset with reasonable confidence.
+
+\subsection{Cluster statistics}
+
+\xxx{Talk about cluster size, quality metric for both datasets}
+
+\subsection{Location statistics}
+
+\xxx{Talk about how many cells we have location data for, the difference between GPS logs and OpenCell, how many cells with $gps_sightings > 1$ and $stdddev < x$ etc.}
+
+\subsection{Effectiveness of clustering algorithm}
+
+In this section we present the most important results of our project: an empirical validation of the clustering algorithm from the point of view of the efficiency of the clusters that it outputs.
+
+\subsubsection{Metrics}
+
+We define two metrics:
+
+\begin{itemize}
+	\item \textbf{Cluster distance}: the average distance between cells in a cluster.
+A good algorithm will group together towers that are geographically close and therefore clusters will have a low distance.
+	\item \textbf{Cluster incompleteness}: the number of cells within $\delta_{incomp}$ meters of another cell $k$ that are not in the same cluster as $k$.
+\end{itemize}
+
+The second metric is necessary, or else the trivial solution of all clusters with size 1 would be optimal.
+
+\subsubsection{Results}
+
+We are only able to assess the cluster distance and incompleteness in the MDC dataset, since it is the only one for which we have GPS information \footnote{Theoretically we could have used OpenCell data to estimate the position of the towers in the RM dataset.
+However, the data does not provide the necessary information about the cell towers, such as network ID.}.
+We choose the MDC GPS logs as our source of location data for the cell towers for two reasons.
+First, as discussed above, there are more cells in the MDC dataset covered by the GPS logs than by the OpenCell database.
+Second, because the GPS logs measure the \textit{user's} location when he or she was connected to that tower, it provides a semantically more accurate value -- after all, the ultimate goal of the clustering algorithm is to provide data that predicts \textit{user's} movements.
+
+\xxx{Talk about\begin{itemize}
+	\item Cluster distance histogram/CDF, average.
+	\item Mean cluster distance/cluster size (no correlation).
+	\item Cluster incompleteness CDF, average.
+	\item What can we conclude from this? That the algorithm works, pretty much.
+\end{itemize}}
 

### 2012-12-12 docs/paper589.tex
Changes to the paper; changed the format to IEEE.

@@
-\documentclass[a4paper,10pt]{article}
+%\documentclass[a4paper,10pt]{article}
+\documentclass[letterpaper, 11pt, conference]{ieeeconf}
 \usepackage{amsmath} % For 'cases' curly brace
@@
 \section{Introduction}
-
+\label{sec:intro}
 \xxx{Talk about mobility prediction in mobile networks. Discuss some of the potential benefits -- some of them were discussed in the
@@
 
+We are only able to assess the cluster distance and incompleteness in the MDC dataset, since it is the only one for which we have GPS information\footnote{In principle we could have used OpenCell data to estimate the position of the towers in the RM dataset.
+However, the data does not provide the necessary information about the cell towers, such as network ID.}.
+We choose the MDC GPS logs as our source of location data for the cell towers for two reasons.
+First, as discussed above, there are more cells in the MDC dataset covered by the GPS logs than by the OpenCell database.
+Second, because the GPS logs measure the \textit{user's} location when he or she was connected to that tower, it provides a semantically more accurate value -- after all, as discussed in Section \ref{sec:intro}, the ultimate goal of the clustering algorithm is to provide data that will be fed to an algorithm that predicts \textit{user's} movements.
+
 \subsection{Dataset comparison}
@@
 
-\xxx{Talk about how many cells we have location data for, the difference between GPS logs and OpenCell, how many cells with $gps_sightings > 1$ and $stdddev < x$ etc.}
+\xxx{Talk about how many cells we have location data for, the difference between GPS logs and OpenCell, how many cells with $gps_{sightings} > 1$ and $stdddev < x$ etc.}
 
@@
 
-We define two metrics:
+We define two metrics, cluster distance and cluster incompleteness.
 
-\begin{itemize}
-	\item \textbf{Cluster distance}: the average distance between cells in a cluster.
+\paragraph{Cluster Distance} the average distance between cells in a cluster.
 A good algorithm will group together towers that are geographically close and therefore clusters will have a low distance.
-	\item \textbf{Cluster incompleteness}: the number of cells within $\delta_{incomp}$ meters of another cell $k$ that are not in the same cluster as $k$.
-\end{itemize}
+The cluster distance is measured in meters.
 
-The second metric is necessary, or else the trivial solution of all clusters with size 1 would be optimal.
-
-\subsubsection{Results}
-
-We are only able to assess the cluster distance and incompleteness in the MDC dataset, since it is the only one for which we have GPS information \footnote{Theoretically we could have used OpenCell data to estimate the position of the towers in the RM dataset.
-However, the data does not provide the necessary information about the cell towers, such as network ID.}.
-We choose the MDC GPS logs as our source of location data for the cell towers for two reasons.
-First, as discussed above, there are more cells in the MDC dataset covered by the GPS logs than by the OpenCell database.
-Second, because the GPS logs measure the \textit{user's} location when he or she was connected to that tower, it provides a semantically more accurate value -- after all, the ultimate goal of the clustering algorithm is to provide data that predicts \textit{user's} movements.
+\paragraph{Cluster incompleteness} For a cell $k$, this is the number of cells within $\delta_{incomp}$ meters of it that are not in the same cluster as $k$.
+The algorithm is shown in \xxx{Show the cluster incompleteness algorithm.} Intuitively this measures if there are potentially oscillating towers that are not being clustered together.
+We can think of the cluster incompleteness as the dual of the cluster distance: the former favors agressive algorithms while the latter encourages conservative approaches.
+Another way to see the benefit of using two metrics is noting that there are trivial solutions that minimize distance or incompleteness (each cell in its own cluster and one huge cluster with all cels, respectively), but there are no trivial solutions that minimize both metrics at the same time.
 
@@
 
+\subsubsection{Cluster distance}
+
+In order to provide some confidence in the precision of the results, we only consider cells that were seen more than once in the GPS logs, with a standard deviation less than 500 meters.
+Only clusters in which more tha 70\% of cells meet this criteria are analyzed.
+This meant that only 82 clusters, or 9.44\% of the total, could be analyzed.
+Although this is a small fraction, there is no reason to think it is a biased sample if we assume there is no correlation between GPS precision and clustering ineffiency.
+Therefore, the results should hold for the entirety of the clusters, and, we conjecture, for other datasets too.
+
+Figures \xxx{cluster distance histogram} and \xxx{cluster distance CDF} show the histogram and CDF of the cluster distances. \xxx{Talk about that.}
+
+We also investigated whether there is a correlation between the size of a cluster and the its distance.
+A graph of both values is presented in Figure \xxx{size x distance}.
+We find that there is no significant correlation.
+
+\subsubsection{Cluster incompleteness}
+
+The measurements for cluster incompleteness are presented in Figures \xxx{cluster incompleteness histogram} and \xxx{cluster incompleteness CDF}.
+The average incompleteness for all cells with GPS data is 0.52 (standard deviation 1.37).
+This is a surprisingly low value and indicates that the algorithm works well.
+
 \section{Conclusion}
 
-\xxx{Talk about other clustering techinques and other modularity measures, like Newman's.}
+We have evaluated a clustering algorithm used to eliminate the oscillation effect in cellular mobility prediction.
+Our main contribution is the use of GPS data to evaluate if the clustered towers are geographically close to each other, which would be circumstantial evidence in favor of the clustering algorithm.
+We find that the clustering algorithm proposed in \cite{mobilityprofiler} performs well on the two metrics we define, cluster distance and cluster incompleteness.
+
+As part of future work, we would like to repeat the experiments here with more data.
+As discussed before, only 9.44\% of the clusters in the MDC dataset had enough information to be reliably processed by the algorithm.
+Another avenue of research is to investigate clustering algorithms that use a different clustering quality metric.
+One possible approach would be the modularity metric defined by Newman and Girvan~\cite{newman}.
+There is also a large body of literature on clustering in theoretical computer science, especially in social networks: Andersen and Peres find sparse cuts efficently in~\cite{sparsecuts}, while Aroroa et.
+al try to find overlapping communities \cite{overlapping}.
+Finding overlapping communities has interesting applications for dealing with the oscillation effect, since towers can conceivably in more than one cluster.
 

### 2012-12-12 src/edu/umich/eecs/statistics/ClusterQualityStatistics.java
Figs are uploaded

@@ -60,16 +60,16 @@ public class ClusterQualityStatistics {
 				}
 				MeanAndStdDev clusterDistances = MeanAndStdDev
 						.fromSample(distances);
-				/*System.out.println(cluster.getCkey().getClusterID() + "\t"
+				System.out.println(cluster.getCkey().getClusterID() + "\t"
 						+ clusterCells.size() + "\t"
 						+ gpsPositionsInCluster.size() + "\t"
 						+ clusterDistances.getMean() + "\t"
-						+ clusterDistances.getStdDev());*/
+						+ clusterDistances.getStdDev());
 				
 				/*double clusterSize=clusterCells.size();
 				double gpsPositionInCluster=gpsPositionsInCluster.size();*/
 				
-				System.out.println(clusterDistances.getStdDev());
+				//System.out.println(clusterDistances.getStdDev());
 						
 			}
 		}

### 2012-12-12 docs/paper589.tex
More changes.

@@
 \maketitle
+
+\xxx{Abstract}
 
@@
 In this project, we analyze one published algorithm for oscillation detection \cite{mobilityprofiler} and evaluate its performance using GPS data as ground-truth.
+We find that this algorithm performs surprisingly well as measured by two metrics we define.
 
-\xxx{``In the next section, we describe....''}
+This report is structured as follows.
+In the next section, we motivate our work by explaining why clustering validation is important.
+In section~\ref{sec:methodology} we describe our methodology  and explain the algorithm we will evaluate.
+We present and discuss results in section~\ref{sec:results}, and conclude in section~\ref{sec:conclusion}.
 \xxx{More about the results.}
@@
 \section{Methodology}
-
+\label{sec:methodology}
 \xxx{What were our methods?\\- Implemented clustering algorithm. Explain mobility paths, oscillations, oscillation graph etc. \\- RM. We sampled the RM dataset and compared it to
@@
 \section{Results}
-
+\label{sec:results}
 \xxx{Results introduction fluff.}
@@
 \section{Conclusion}
-
+\label{sec:conclusion}
 We have evaluated a clustering algorithm used to eliminate the oscillation effect in cellular mobility prediction.
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{IEEEtranS}
 \bibliography{eecs589Report}

### 2012-12-12 docs/paper589.tex
Abstract.

@@
 
-\xxx{Abstract}
+\begin{abstract}
+We investigate in detail one of the algorithms involved in cellular mobility prediction.
+These systems try to predict a user's position based on cell tower connections.
+One key difficulty in this is the oscillation effect: a stationary phone will often switch between towers, adding significant noise to the data.
+We look at the efficiency of one of the approaches in the literature that tries to counter this effect.
+This algorithm tries to cluster cell towers that users often oscillate between.
+In order to do evaluate the algorithm's efficiency, we use a new dataset that allows us to infer GPS data for cell towers.
+We find that the algorithm performs well.
+\end{abstract}
 
@@
 We present and discuss results in section~\ref{sec:results}, and conclude in section~\ref{sec:conclusion}.
-\xxx{More about the results.}
 

### 2012-12-12 docs/paper589.tex
typo

@@
 In this project, we address one of the challenges faced in mobility prediction using GSM cell tower connection logs. In our context, mobility prediction in cellular
-towers can be defined as the ability to predict a user's physical movement by examning the history of cell tower connections. This is very important
+towers can be defined as the ability to predict a user's physical movement by examining the history of cell tower connections. This is very important
 for mobile network operators, as several network optimizations are possible if they can forecast with reasonable accuracy

### 2012-12-12 docs/paper589.tex
typo

@@
 \cite{mobilityprofiler}, and the Leap Graph paper by Duffield et al \cite{LeapGraph}.
-Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by resarchers at AT\&T, makes use of the active set of connections.
+Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by researchers at AT\&T, makes use of the active set of connections.
 The active set contains all towers that are within range of the phone's antenna at any given time.

### 2012-12-12 docs/paper589.tex
typo

@@
 The example in \cite{mobilityprofiler} is illustrative: consider a sequence of handovers $\left[x, y, x, w, v, w, y\right]$ where every element represents one cell tower.
-We say that the pair $(x, y)$ has 3 switches: the first twofrom $x$ to $y$ and back to $x$ in the first positions of the sequence, and then back to $y$ at the end.
+We say that the pair $(x, y)$ has 3 switches: the first two from $x$ to $y$ and back to $x$ in the first positions of the sequence, and then back to $y$ at the end.
 Hence, if we adopt $k=3$ the only oscillating pair would be $(x, y)$.
@@
 One possible approach would be the modularity metric defined by Newman and Girvan~\cite{newman}.
-There is also a large body of literature on clustering in theoretical computer science, especially in social networks: Andersen and Peres find sparse cuts efficently in~\cite{sparsecuts}, while Aroroa et.
+There is also a large body of literature on clustering in theoretical computer science, especially in social networks: Andersen and Peres find sparse cuts efficiently in~\cite{sparsecuts}, while Aroroa et.
 al try to find overlapping communities \cite{overlapping}.

### 2012-12-12 docs/paper589.tex
typo

@@
 
-\xxx{Compare MDC and Reality Mining, i.e. show that they are similiar using the sampled RM?}
+\xxx{Compare MDC and Reality Mining, i.e. show that they are similar using the sampled RM?}
 
@@
 The algorithm is shown in \xxx{Show the cluster incompleteness algorithm.} Intuitively this measures if there are potentially oscillating towers that are not being clustered together.
-We can think of the cluster incompleteness as the dual of the cluster distance: the former favors agressive algorithms while the latter encourages conservative approaches.
+We can think of the cluster incompleteness as the dual of the cluster distance: the former favors aggressive algorithms while the latter encourages conservative approaches.
 Another way to see the benefit of using two metrics is noting that there are trivial solutions that minimize distance or incompleteness (each cell in its own cluster and one huge cluster with all cels, respectively), but there are no trivial solutions that minimize both metrics at the same time.
@@
 This meant that only 82 clusters, or 9.44\% of the total, could be analyzed.
-Although this is a small fraction, there is no reason to think it is a biased sample if we assume there is no correlation between GPS precision and clustering ineffiency.
+Although this is a small fraction, there is no reason to think it is a biased sample if we assume there is no correlation between GPS precision and clustering inefficiency.
 Therefore, the results should hold for the entirety of the clusters, and, we conjecture, for other datasets too.

### 2012-12-12 docs/paper589.tex
Stationary oscillation example.

--- 2012-12-12 docs/paper589.tex
@@
 \usepackage{url}
+\usepackage{graphicx}
+\usepackage[font=bf]{caption}
 
@@
 \label{sec:intro}
-\xxx{Talk about mobility prediction in mobile networks. Discuss some of the potential benefits -- some of them were discussed in the
-ATT leap graph paper. Maybe do a little research on this.}
 
 In this project, we address one of the challenges faced in mobility prediction using GSM cell tower connection logs. In our context, mobility prediction in cellular
-towers can be defined as the ability to predict a user's physical movement by examining the history of cell tower connections. This is very important
+towers can be defined as the ability to predict a user's physical movement by examning the history of cell tower connections. This is very important
 for mobile network operators, as several network optimizations are possible if they can forecast with reasonable accuracy
@@
 A major challenge in this is dealing with the \textit{oscillation effect}, which happens when a \textit{stationary} user's phone switches between two or more towers.
+Figure~\ref{fig:stationary} shows an example of a stationary phone that connects to different cell towers.
 This is a natural consequence of the radio medium and the fact that towers very often have overlapping ranges.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.15\textwidth]{figs/stationary_oscillation}
+\caption{An example of actual cell tower connection logs for a stationary phone, slightly modified from~\cite{LeapGraph}.
+The time of the sample is on the left, and an identifier for the cell tower is on the right.
+We can see that the phone connects to several towers, even though it does not move.
+}
+\label{fig:stationary}
+\end{figure}
+
 \section{Motivation}
@@
 \xxx{How do we cite the Nokia MDC?}
-
-\xxx{Reproduce the stationary phone example in the leap graph paper.}
 
@@
 \cite{mobilityprofiler}, and the Leap Graph paper by Duffield et al \cite{LeapGraph}.
-Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by researchers at AT\&T, makes use of the active set of connections.
+Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by resarchers at AT\&T, makes use of the active set of connections.
 The active set contains all towers that are within range of the phone's antenna at any given time.

### 2012-12-13 docs/paper589.tex
opencell

--- 2012-12-12 docs/paper589.tex
@@
 \label{sec:intro}
-\xxx{Talk about mobility prediction in mobile networks. Discuss some of the potential benefits -- some of them were discussed in the
-ATT leap graph paper. Maybe do a little research on this.}
 
@@
 
-\subsubsection{OpenCell}
+\subsubsection{OpenCellId and OpenBMap}
+
+OpenCellId and OpenBmap are two open source projects with goal of providing researchers with a complete database of  cells with their locations.
+We use these two dataset for retrieving location information of cells
+we have in MDC.
 

### 2012-12-13 docs/paper589.tex
First draft of the paper.

@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[letterpaper, 11pt, conference]{ieeeconf}
+\documentclass[letterpaper, 12pt, conference]{ieeeconf}
 \usepackage{amsmath} % For 'cases' curly brace
@@
 \usepackage{amssymb}
+\usepackage{array}
 \usepackage{url}
 \usepackage{graphicx}
+\usepackage{subcaption}
+
 \usepackage[font=bf]{caption}
@@
 \begin{document}
-\title{Validating Cell Tower Clustering for Mobility Prediction}
-\author{Mehrdad Moradi\\moradi@umich.edu \and Pedro d'Aquino\\pdaquino@umich.edu}
+\title{\LARGE \bf
+Validating Cell Tower Clustering for Mobility Prediction
+}
+\author{Mehrdad Moradi\\moradi@umich.edu \and Pedro d'Aquino\\
+pdaquino@umich.edu}
 \date{December 11, 2012}
@@
 \begin{abstract}
-We investigate in detail one of the algorithms involved in cellular mobility prediction.
-These systems try to predict a user's position based on cell tower connections.
-One key difficulty in this is the oscillation effect: a stationary phone will often switch between towers, adding significant noise to the data.
-We look at the efficiency of one of the approaches in the literature that tries to counter this effect.
-This algorithm tries to cluster cell towers that users often oscillate between.
-In order to do evaluate the algorithm's efficiency, we use a new dataset that allows us to infer GPS data for cell towers.
-We find that the algorithm performs well.
+We investigate in detail one of the algorithms involved in cellular mobility 
+prediction. These systems try to predict a user's position based on cell 
+tower connections. One key difficulty in this is the oscillation effect: a 
+stationary phone will often switch between towers, adding significant noise 
+to the data. We look at the efficiency of one of the approaches in the 
+literature that tries to counter this effect.  This algorithm tries to 
+cluster cell towers that users often oscillate between. In order to do 
+evaluate the algorithm's efficiency, we use a new dataset that allows us to 
+infer GPS data for cell towers. We find that the algorithm performs well.
 \end{abstract}
+\xxx{Make sure I didnt write Mobility Data Challenge (it should be Mobile 
+data challenge), or OpenCell (it should be open dataset).}
 
@@
 
-In this project, we address one of the challenges faced in mobility prediction using GSM cell tower connection logs. In our context, mobility prediction in cellular
-towers can be defined as the ability to predict a user's physical movement by examining the history of cell tower connections. This is very important
-for mobile network operators, as several network optimizations are possible if they can forecast with reasonable accuracy
-where a user will be in the future \cite{LeapGraph} \cite{Liu97anoptimal}. By being able to pre-allocate resources to towers that are most likely to
-be connected to by the user, network operators can improve latency and uptime. This is increasingly important as the amount of data in cellular networks
-increases: indeed, in 2010 the amount of data traffic surpassed voice traffic for the first time, and it is expected to double annually in the near future
+In this project, we address one of the challenges faced in mobility 
+prediction using GSM cell tower connection logs. In our context, mobility 
+prediction in cellular
+towers can be defined as the ability to predict a user's physical movement by 
+examining the history of cell tower connections. This is very important
+for mobile network operators, as several network optimizations are possible 
+if they can forecast with reasonable accuracy
+where a user will be in the future \cite{LeapGraph} \cite{Liu97anoptimal}. By 
+being able to pre-allocate resources to towers that are most likely to
+be connected to by the user, network operators can improve latency and 
+uptime. This is increasingly important as the amount of data in cellular 
+networks
+increases: indeed, in 2010 the amount of data traffic surpassed voice traffic 
+for the first time, and it is expected to double annually in the near future
 \cite{EricssonData}.
@@
 Another use of mobility prediction is to discover mobility profiles, i.e. 
-patterns in a user's movement \cite{mobilityprofiler}.
-As illustrated in \cite{mobilityprofiler}, mobility profiles have a wide range of possible applications, from traffic planning to advertisement.
-Recently, Google started integrating
-a contextual search and recommendation service in its phones that is able to, for instance, detect a user's home and work addresses and suggest routes based on
+patterns in a user's movement \cite{mobilityprofiler}. As illustrated in \cite
+{mobilityprofiler}, mobility profiles have a wide range of possible 
+applications, from traffic planning to advertisement. Recently, Google 
+started integrating
+a contextual search and recommendation service in its phones that is able to, 
+for instance, detect a user's home and work addresses and suggest routes 
+based on
 traffic \cite{googleNow}.
 
-The heart of every mobility prediction system based on cell tower connections is to infer movement based on which cell the user's phone is connected to. The goal
-is to predict which tower the user is more likely to connect to next.
-A major challenge in this is dealing with the \textit{oscillation effect}, which happens when a \textit{stationary} user's phone switches between two or more towers.
-Figure~\ref{fig:stationary} shows an example of a stationary phone that connects to different cell towers.
-This is a natural consequence of the radio medium and the fact that towers very often have overlapping ranges.
-In order to accurately predict a user's mobility, these outliers have to be removed so that fluctuations can be separated from actual user movement.
+The heart of every mobility prediction system based on cell tower connections 
+is to infer movement based on which cell the user's phone is connected to. 
+The goal
+is to predict which tower the user is more likely to connect to next. A major 
+challenge in this is dealing with the \textit{oscillation effect}, which 
+happens when a \textit{stationary} user's phone switches between two or more 
+towers. Figure~\ref{fig:stationary} shows an example of a stationary phone 
+that connects to different cell towers. This is a natural consequence of the 
+radio medium and the fact that towers very often have overlapping ranges. In 
+order to accurately predict a user's mobility, these outliers have to be 
+removed so that fluctuations can be separated from actual user movement.
 
-In this project, we analyze one published algorithm for oscillation detection \cite{mobilityprofiler} and evaluate its performance using GPS data as ground-truth.
-We find that this algorithm performs surprisingly well as measured by two metrics we define.
+In this project, we analyze one published algorithm for oscillation detection 
+\cite{mobilityprofiler} and evaluate its performance using GPS data as ground-
+truth. We find that this algorithm performs surprisingly well as measured by 
+two metrics we define.
 
-This report is structured as follows.
-In the next section, we motivate our work by explaining why clustering validation is important.
-In section~\ref{sec:methodology} we describe our methodology  and explain the algorithm we will evaluate.
-We present and discuss results in section~\ref{sec:results}, and conclude in section~\ref{sec:conclusion}.
+This report is structured as follows. In the next section, we motivate our 
+work by explaining why clustering validation is important.
+In section~\ref{sec:methodology} we describe our methodology  and explain the algorithm we 
+will evaluate. We present and discuss results in section~\ref{sec:results}, 
+and conclude in section~\ref{sec:conclusion}.
 
@@
 \includegraphics[width=0.15\textwidth]{figs/stationary_oscillation}
-\caption{An example of actual cell tower connection logs for a stationary phone, slightly modified from~\cite{LeapGraph}.
-The time of the sample is on the left, and an identifier for the cell tower is on the right.
-We can see that the phone connects to several towers, even though it does not move.
-}
+\caption{An example of actual cell tower connection logs for a stationary 
+phone, slightly modified from~\cite{LeapGraph}. The time of the sample is on 
+the left, and an identifier for the cell tower is on the right. We can see 
+that the phone connects to several towers, even though it does not move. }
 \label{fig:stationary}
@@
 
-\xxx{This is where we explain why we think that our work is important: because current tower clustering methods do not have ground truth data to check against and we do.}
+The accuracy of current mobility prediction systems based on cell tower 
+connections is usually measured through machine learning techniques like cross
+-validation that try to validate the ultimate goal of forecasting the user's 
+future connections \cite{LeapGraph}. As far as we know there are no studies 
+that evaluate methods of dealing with the oscillation effect, even though all 
+prediction systems must deal with it.
 
-\xxx{How do we cite the Nokia MDC?}
-
-The accuracy of current mobility prediction systems based on cell tower connections is usually measured through machine learning techniques like cross-validation that try to validate the ultimate goal of forecasting the user's future connections \cite{LeapGraph}.
-As far as we know there are no studies that evaluate methods of dealing with the oscillation effect, even though all prediction systems must deal with it.
-
-In this paper we report the results of such an analysis.
-We have implemented and evaluated the ``oscillation graph'' technique described in \cite{mobilityprofiler}.
-The basic idea of this algorithm is to detect cell towers that often present oscillations and consider them as a single tower.
-This is effectively a clustering algorithm.
-Our goal is to evaluate whether towers that are clustered together are actually close in real life.
-To evaluate this, we use the Nokia Mobility Challenge dataset, which has data on both cell tower connection (``GSM logs'') and the GPS position of the phones.
+In this paper we report the results of such an analysis. We have implemented 
+and evaluated the ``oscillation graph'' technique described in \cite{
+mobilityprofiler}. The basic idea of this algorithm is to detect cell towers 
+that often present oscillations and consider them as a single tower. This is 
+effectively a clustering algorithm. Our goal is to evaluate whether towers 
+that are clustered together are actually close in real life. To evaluate 
+this, we use the Nokia Mobility Challenge dataset, which has data on both 
+cell tower connection (``GSM logs'') and the GPS position of the phones.
 
@@
 \label{sec:methodology}
-\xxx{What were our methods?\\- Implemented clustering algorithm. Explain mobility paths, oscillations, oscillation graph etc. \\- RM. We sampled the RM dataset and compared it to
-the Nokia MDC \\- MDC. We apply the oscillation clustering to the MDC and look at the position
-characteristics of the clusters.}
 
-The Mobility Profiler paper \cite{mobilityprofiler} describes an algorithm to detect the oscillation effect.
-Our approach to evaluating this algorithm's efficiency was the following.
-We implement the algorithm as described in \cite{mobilityprofiler} and test it in the same dataset used in the original paper.
-However, that dataset does not have any data on the location of the users when they were connected to the towers.
-We therefore apply the algorithm to another dataset, from the Nokia Mobility Challenge, in which the GPS position of the users is listed.
-In order to evaluate how successful the clustering algorithm is, we devise two metrics: cluster distance and cluster incompleteness.
+The Mobility Profiler paper \cite{mobilityprofiler} describes an algorithm to 
+detect the oscillation effect. Our approach to evaluating this algorithm's 
+efficiency was the following. We implement the algorithm as described in \cite
+{mobilityprofiler} and test it in the same dataset used in the original 
+paper. However, that dataset does not have any data on the location of the 
+users when they were connected to the towers. We therefore apply the 
+algorithm to another dataset, from the Nokia Mobility Challenge, in which the 
+GPS position of the users is listed. In order to evaluate how successful the 
+clustering algorithm is, we devise two metrics: cluster distance and cluster 
+incompleteness.
 
@@
 
-There are two main papers that tackle the problem of oscillation effects: the Mobility Profiler paper, by Bayir et al.
-\cite{mobilityprofiler}, and the Leap Graph paper by Duffield et al \cite{LeapGraph}.
-Ideally we would want to evaluate both of them, but the leap graph algorithm, which was published by researchers at AT\&T, makes use of the active set of connections.
-The active set contains all towers that are within range of the phone's antenna at any given time.
-Unfortunately, active set logs are not public which makes it impossible for us to evaluate the leap graph algorithm.
+There are two main papers that tackle the problem of oscillation effects: the 
+Mobility Profiler paper, by Bayir et al. \cite{mobilityprofiler}, and the 
+Leap Graph paper by Duffield et al \cite{LeapGraph}. Ideally we would want to 
+evaluate both of them, but the leap graph algorithm, which was published by 
+researchers at AT\&T, makes use of the active set of connections. The active 
+set contains all towers that are within range of the phone's antenna at any 
+given time. Unfortunately, active set logs are not public which makes it 
+impossible for us to evaluate the leap graph algorithm.
 
-We therefore focus our efforts on the algorithm described in the Mobility Profiler paper, which we call the oscillation graph algorithm.
-In the interest of completeness, we briefly describe how it works.
+We therefore focus our efforts on the algorithm described in the Mobility 
+Profiler paper, which we call the oscillation graph algorithm. In the 
+interest of completeness, we briefly describe how it works.
 
-\xxx{Cellspan log example. Mobility path example.}
+\begin{figure*}
+        \centering
+        \begin{subfigure}[t]{0.44\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{figs/gsm_log_rm}
+                \caption{A series of timestamped GSM cell tower connection 
+log records, from the Reality Mining dataset. The cell tower is identified by 
+an area code and a cell ID, which are separated by a period. In this 
+particular example it seems that all three towers are oscillating.}
+                \label{fig:gsm_log_example}
+        \end{subfigure}%
+        \qquad %add desired spacing between images, e. g. ~, \quad, \qquad etc. 
+          %(or a blank line to force the subfigure onto a new line)
+        \begin{subfigure}[t]{0.48\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{figs/gsm_log_mdc}
+                \caption{An example of GSM records in the Nokia Mobile Data 
+Challenge. From left to right, the fields are: user id, timestamp, time zone, 
+country ID, network ID, cell ID, area ID, signal strength from 1 to 7 and 
+signal strength in dBm.}
+			\label{fig:gsm_log_mdc}
+        \end{subfigure}
+				\caption{Example of GSM cell tower connection logs in both datasets 
+we use.}
+\end{figure*}
 
 \subsubsection{Mobility paths}
-The input of the algorithm is a sequence of \textit{handovers}, which are transition events when a phone switches the tower it is connected to.
-These events are timestamped.
-The first step of the algorithm is to analyze the transitions and extract \textit{mobility paths}.
-A mobility path corresponds to a sequence of handovers that are associated with actual user movement.
-For example, imagine a user that drives every morning to work, stays there for eight hours and goes back home.
-During his drive to work, the user's phone will switch cell towers several times.
-These handovers will be included in a mobility path.
-While the user is at work, ignoring the oscillation effect temporarily, the phone will be connected to a single tower.
-Finally, when the user the goes home the handovers that occur during the drive will belong to a different mobility path.
+The input of the algorithm is a sequence of \textit{handovers}, which are 
+transition events when a phone switches the tower it is connected to. These 
+events are timestamped, as shown in Figure~\ref{fig:gsm_log_example}. The 
+first step of the algorithm is to analyze the transitions and extract \textit{
+mobility paths}.  A mobility path corresponds to a sequence of handovers that 
+are associated with actual user movement. For example, imagine a user that 
+drives every morning to work, stays there for eight hours and goes back home. 
+During his drive to work, the user's phone will switch cell towers several 
+times. These handovers will be included in a mobility path. While the user is 
+at work, ignoring the oscillation effect temporarily, the phone will be 
+connected to a single tower. Finally, when the user the goes home the 
+handovers that occur during the drive will belong to a different mobility path.
 
-More formally, let $\Delta_i$ be the time difference between two consecutive handovers for a particular user's connection log.
-In other words, $\Delta_i$ is the time the user spent connected to cell tower $C_i$.
-Let $\delta_{duration}$ be the \textit{end-location threshold}.
-Then we define a mobility path as a sequence of cell connections $\left\{C_i, C_{i+1}, \ldots, C_j\right\}$ such that $\Delta_i < \delta_{duration}$ for all handovers.\footnote{We also have to consider \textit{hidden end-locations}, which happen when a user's phone is disconnected from any towers, or if the user turns his phone off.
-We omit its formal definition for briefness.}
+More formally, let $\Delta_i$ be the time difference between two consecutive 
+handovers for a particular user's connection log. In other words, $\Delta_i$ 
+is the time the user spent connected to cell tower $C_i$. Let $\delta_{
+duration}$ be the \textit{end-location threshold}. Then we define a mobility 
+path as a sequence of cell connections $\left\{C_i, C_{i+1}, \ldots, C_j
+\right\}$ such that $\Delta_i < \delta_{duration}$ for all handovers.\footnote
+{We also have to consider \textit{hidden end-locations}, which happen when a 
+user's phone is disconnected from any towers, or if the user turns his phone 
+off. We omit its formal definition for briefness.}
 
@@
 
-The goal of the oscillation detection algorithm is to retrieve sets of towers that overlap each other -- that is, towers that could be in a stationary user's handover logs.
-The main idea behind the algorithm is that a user's phone is more likely to switch between oscillating towers than between regular towers that are not overlapping.
-In this context, \textit{switching} means a handover from one tower to another inside of a mobility path.
-We can define $k$ to be the minimum switching count and consider all pairs of towers that have more than $k$ switches in one mobility path to be an \textit{oscillating pair}.
-The example in \cite{mobilityprofiler} is illustrative: consider a sequence of handovers $\left[x, y, x, w, v, w, y\right]$ where every element represents one cell tower.
-We say that the pair $(x, y)$ has 3 switches: the first two from $x$ to $y$ and back to $x$ in the first positions of the sequence, and then back to $y$ at the end.
-Hence, if we adopt $k=3$ the only oscillating pair would be $(x, y)$.
-Note that switches do not need to be consecutive.
-The authors argue that this allows for detection of overlapping towers in dense region where multiple towers may be within range at the same time.
+The goal of the oscillation detection algorithm is to retrieve sets of towers 
+that overlap each other -- that is, towers that could be in a stationary 
+user's handover logs. The main idea behind the algorithm is that a user's 
+phone is more likely to switch between oscillating towers than between 
+regular towers that are not overlapping. In this context, \textit{switching} 
+means a handover from one tower to another inside of a mobility path. We can 
+define $\delta_{osc}$ to be the minimum switching count and consider all 
+pairs of towers that have more than $\delta_{osc}$ switches in one mobility 
+path to be an \textit{oscillating pair}. The example in \cite{mobilityprofiler}
+is illustrative: consider a sequence of handovers $\left[x, y, x, w, v, w, y
+\right]$ where every element represents one cell tower. We say that the pair $
+(x, y)$ has 3 switches: the first two from $x$ to $y$ and back to $x$ in the 
+first positions of the sequence, and then back to $y$ at the end. Hence, if 
+we adopt $\delta_{osc}=3$ the only oscillating pair would be $(x, y)$. Note 
+that switches do not need to be consecutive. The authors argue that this 
+allows for detection of overlapping towers in dense region where multiple 
+towers may be within range at the same time.
 
@@
 
-After detecting all the oscillating pairs in all mobility paths, the next step is to build the oscillation graph.
-The set $V$ of vertices in the graph are all cell towers in the dataset, and the weighted edges in $E$ connect oscillating pairs.
-Let $P_i$ be a mobility path.The weight of every edge, which we also call its \textit{support}, is defined as:
+After detecting all the oscillating pairs in all mobility paths, the next 
+step is to build the oscillation graph. The set $V$ of vertices in the graph 
+are all cell towers in the dataset, and the weighted edges in $E$ connect 
+oscillating pairs. Let $P_i$ be a mobility path.The weight of every edge, 
+which we also call its \textit{support}, is defined as:
 
 \begin{equation*}
-s(x,y) = \frac{|\left\{P_i| (x, y) \in P_i \wedge (x, y) \text{oscillated}\right\}|}{|\left\{P_i| (x, y) \in P_i\right\}|}
+s(x,y) = \frac{|\left\{P_i| (x, y) \in P_i \wedge (x, y) \text{oscillated}
+\right\}|}{|\left\{P_i| (x, y) \in P_i\right\}|}
 \end{equation*}
 
-In words, $s(x,y)$ is the ratio of how many times the cell towers $x$ and $y$ oscillated over the number of mobility paths in which $x$ and $y$ appear.
-For example, a pair of towers such that in \textit{every} mobility path that both of them appear the switch count is greater than $k$ would have a support equal to 1.
-If a pair only switches more or equal to $k$ times in half the mobility paths both of the towers appear in, that pair's support would be 0.5.
+In words, $s(x,y)$ is the ratio of how many times the cell towers $x$ and $y$ 
+oscillated over the number of mobility paths in which $x$ and $y$ appear. For 
+example, a pair of towers such that in \textit{every} mobility path that both 
+of them appear the switch count is greater than $k$ would have a support 
+equal to 1. If a pair only switches more or equal to $k$ times in half the 
+mobility paths both of the towers appear in, that pair's support would be 0.5.
 
@@
 
-\xxx{Explain the clustering algorithm.}
-
 The final step of the algorithm is to find clusters in the oscillation graph. 
-Intuitively, a cluster represents a group of towers that are not indicative of actual user movement.
-The Mobility Profiler paper uses an greedy divise algorithm for clustering.
-At every iteration, the algorithm removes the edge with the lowest weight from the graph.
-It then evaluates the ``quality'' of all connected components, which is the weighted ratio of edges inside of the cluster over the number of edges that leave the cluster in the original graph.
-Formally, let $C$ be a cluster with the set of edges $E$, and $E_{out}$ be the set of edges that have exactly one vertex in $C$, and define $w(e)$ to be an edge's weight:
+Intuitively, a cluster represents a group of towers that are not indicative 
+of actual user movement. The Mobility Profiler paper uses an greedy divise 
+algorithm for clustering. At every iteration, the algorithm removes the edge 
+with the lowest weight from the graph. It then evaluates the ``quality'' of 
+all connected components, which is the weighted ratio of edges inside of the 
+cluster over the number of edges that leave the cluster in the original 
+graph. Formally, let $C$ be a cluster with the set of edges $E$, and $E_{out}$
+ be the set of edges that have exactly one vertex in $C$, and define $w(e)$ 
+to be an edge's weight:
 
 \begin{equation*}
-Q(C) = \frac{\displaystyle\sum_{\forall e_{in} \in E}{w(e_{in})}}{\displaystyle\sum_{\forall e_{out} \in E_{out}}{w(e_{out})}}
+Q(C) = \frac{\displaystyle\sum_{\forall e_{in} \in E}{w(e_{in})}}{
+\displaystyle\sum_{\forall e_{out} \in E_{out}}{w(e_{out})}}
 \end{equation*}
 
-Every cluster with quality metric greater than a threshold is removed from the graph.
-The algorithm continues until there are no more vertices in the graph.
-The algorithm outputs all the clusters it found.
-Our main goal is to evaluate these clusters according to the GPS position of the users when they were within range of that tower.
+Every cluster with quality metric greater than a threshold is removed from 
+the graph. The algorithm continues until there are no more vertices in the 
+graph. The algorithm outputs all the clusters it found. Our main goal is to 
+evaluate these clusters according to the GPS position of the users when they 
+were within range of that tower.
 
@@
 
-\xxx{Reality Mining dataset reference.}
-
-We use two datasets with GSM cell tower connection data, from the Reality Mining project and from the Nokia Mobility Data Challenge.
+We use two datasets with GSM cell tower connection data, from the Reality 
+Mining project and from the Nokia Mobile Data Challenge.
 
@@
 
-\xxx{Describe the reality mining dataset}
+The Reality Mining dataset (RM)~\cite{rm} was obtained from over 330 thousand 
+hours of data logged from 94 volunteers' cell phones during 2004. The breadth 
+of information collected is impressive, ranging from cell tower IDs to which 
+Bluetooth devices are within 5 meters of the users. This dataset has been 
+widely used in research into social and friendship networks. For our project, 
+following~\cite{mobilityprofiler}, the focus is the cell tower connection logs.
 
-\subsubsection{Nokia Mobility Data Challenge -- GSM}
+\subsubsection{Nokia Mobile Data Challenge -- GSM}
 
-\xxx{Describe the Nokia Mobility Challenge}
+The Nokia Mobile Data Challenge (MDC) project~\cite{nokiaMdc} gathered data 
+from 170 volunteers in the Swiss city of Lausanne during 2009. In comparison 
+to the Reality Mining project, more sensors were available for the 
+researchers (a consequence of the evolution of mobile phones) and the 
+population was more diverse (two thirds of the volunteers were employed). In 
+particular, there is data available on the users' cell tower connections, 
+which is essential for mobility prediction. However, while the Reality Mining 
+dataset logs every cell transition, the MDC authors sample the active cell 
+tower connection every minute. This leads to some data loss, since cell 
+transitions can happen at a much higher frequency (e.g.
+Figure~\ref{fig:gsm_log_example}). An example of the GSM logs in the MDC is shown in
+Figure~\ref{fig:gsm_log_mdc}.
 
@@
 
-\xxx{GPS logs from MDC and OpenCell}
-
-Our main contribution is an analysis of the oscillation clustering algorithm using GPS data. We use two sources of GPS data, the Nokia MDC and the OpenCell database.
+Our main contribution is an analysis of the oscillation clustering algorithm 
+using GPS data. We use two sources of GPS data, the Nokia MDC and the 
+OpenCell database.
 
@@
 
-Phones in the MDC also had their GPS units sampled every minute.
-The format of the GPS logs is shown in Figure \xxx{Show the MDC GPS logs}.
-We process the GPS and try to establish an approximate position for each cell tower.
-We approximate the cell tower position as the average of the positions of users when they were connected to that tower.
-To do this, we have to match each entry in the GSM log with an entry in the GPS log.
+Phones in the MDC also had their GPS units sampled every minute, recording 
+the latitude and longitude of the user's phone. We process the GPS and try to 
+establish an approximate position for each cell tower. We approximate the 
+cell tower position as the average of the positions of users when they were 
+connected to that tower. To do this, we have to match each entry in the GSM 
+log with an entry in the GPS log.
 
-Because GSM and GPS entries do not necessarily have the same timestamp (in fact, they almost never do), we use GPS entries that have a timestamp within $\delta_{GPS}$ seconds of the GSM entry we are trying to match.
-The algorithm goes through every record in the GSM log and tries to find a matching GPS log.
-After going through all users' connection logs, we estimate each tower's position as the average of all GPS positions that were associated with that tower.
-It is important to have a measure of how accurate the GPS estimate of a particular tower is, so we record the number of GPS samples for that tower and the standard deviation in the samples.
-The standard deviation is computed based on the distance from each sample to the average point, which is calculated using the Haversine formula \cite{haversine}.
+Because GSM and GPS entries do not necessarily have the same timestamp (in 
+fact, they almost never do), we use GPS entries that have a timestamp within $
+\delta_{GPS}$ seconds of the GSM entry we are trying to match. The algorithm 
+goes through every record in the GSM log and tries to find a matching GPS 
+log. After going through all users' connection logs, we estimate each tower's 
+position as the average of all GPS positions that were associated with that 
+tower. It is important to have a measure of how accurate the GPS estimate of 
+a particular tower is, so we record the number of GPS samples for that tower 
+and the standard deviation in the samples. The standard deviation is computed 
+based on the distance from each sample to the average point, which is 
+calculated using the Haversine formula \cite{haversine}. In the experiments in this
+report we used $\delta_{GPS}=40$~seconds.
 
-The algorithm runs in $O(k(n\log n + m\log n))$, where $k$ is the number of users in the dataset and $m$ and $n$ are upper bounds on the number of GPS and GSM records per user, respectively.
-The logarithmic factor comes from sorting the GPS log by timestamp and performing a binary search.
+The algorithm runs in $O(k(n\log n + m\log n))$, where $k$ is the number of 
+users in the dataset and $m$ and $n$ are upper bounds on the number of GPS 
+and GSM records per user, respectively. The logarithmic factor comes from 
+sorting the GPS log by timestamp and performing a binary search.
 
-\subsubsection{OpenCellId and OpenBMap}
+\subsubsection{OpenCellId and OpenBMap (open dataset)}
 
-OpenCellId and OpenBmap are two open source projects with goal of providing researchers with a complete database of  cells with their locations.
-We use these two dataset for retrieving location information of cells
-we have in MDC.
-
-\xxx{Talk about OpenCell; what it is and how we use it.}
+OpenCellID~\cite{openCellId} and OpenBmap~\cite{openBMap} are two open source 
+projects with the goal of providing researchers with a complete database of 
+cells towers and their locations They gather the data through users who 
+voluntarily participate in the project by adding a logging component to their 
+phones that keeps track of the position and associates it with a cell tower. 
+OpenBMap has over 791694 cells in its database, while OpenCellID covers 
+1889463. In our application we combine both measures and use an average, 
+which we refert to as the ``open dataset''.
 
@@
 \label{sec:results}
-\xxx{Results introduction fluff.}
 
-\xxx{Talk about challenges in handling the data -- wrong clock, entries out of order in the RM dataset;}
+\begin{table*}
+    \centering
+    \begin{tabular}{ | c || r | r | r | r | r | }
+    \hline
+    Dataset & Cells  & Area codes & Country IDs & Handovers & Clusters \\
+    \hline
+    Nokia Mobile Data Challenge & 23378 & 966 & 33 & 3308612 & 869 \\
+    \hline
+    Reality Mining & 32787 & 32787 & 2 & 32787 & 32787 \\
+    \hline
+    \end{tabular}
+\caption{Comparison between the Nokia Mobile Data and Reality Mining datasets.}
+\label{tbl:datasetComparison}
+\end{table*}
 
-We implemented the clustering algorithm in \cite{mobilityprofiler} for the Reality Mining and the Nokia Mobility Data Challenge datasets and compared the results of the clustering in both datasets (adjusting for the fact that the MDC dataset's data is sampled every minute).
-In this section, we report the results of running the algorithm in those datasets and analyze the results.
-\xxx{One sentence description of the non-GPS related results.} We also used GPS data to evaluate the accuracy and quality of the resulting clusters.
-\xxx{One sentence description of the GPS results}.
+We implemented the clustering algorithm in \cite{mobilityprofiler} for the 
+Reality Mining and the Nokia Mobility Data Challenge datasets and compared 
+the results of the clustering in both datasets (adjusting for the fact that 
+the MDC dataset's data is sampled every minute). In this section, we report 
+the results of running the algorithm in those datasets and analyze the 
+results. We find that the Reality Mining and MDC datasets have similar 
+patterns of cluster size and density. We also used GPS data to evaluate the 
+accuracy and quality of the resulting clusters. Our experiments show that the 
+clustering algorithm performs well on both metrics we define.
 
-Running the clustering algorithm on the RM dataset was more challenging than originally expected; there were inconsistencies with the dates (apparently the phones' clocks reverted to 01/01/2004 when the battery died) and time (with entries appearing in an inconsistent order; see \xxx{RM GSM time weirdness example}).
-We dealt with these problems by ignoring records with a timestamp equal to 01/01/2004 and by assuming that the ordering of the lines in the dataset logs represented the actual ordering that the logs were collected -- this seemed consistent with manual analysis of the logs and generated valid mobility paths, which is not the case if the timestamps are taken at face value.
-\xxx{Maybe expand on what a valid mobility path is.}
+Running the clustering algorithm on the RM dataset was more challenging than 
+originally expected; there were inconsistencies with the dates (apparently 
+the phones' clocks reverted to 01/01/2004 when the battery died) and time (
+with entries appearing in an inconsistent order. We dealt with these problems 
+by ignoring records with a timestamp equal to 01/01/2004 and by assuming that 
+the ordering of the lines in the dataset logs represented the actual ordering 
+that the logs were collected -- this seemed consistent with manual analysis 
+of the logs and generated valid mobility paths\footnote{A valid mobility path 
+is one in which there is no overlap of handovers -- that is, there cannot be 
+two cells connected at the same time.}, which is not the case if the 
+timestamps are taken at face value.
 
-In all experiments we used the same parameters value as the original Mobility Profiler paper, namely \xxx{list the constants we used, like the minimum oscillation count etc.}.
-
-We are only able to assess the cluster distance and incompleteness in the MDC dataset, since it is the only one for which we have GPS information\footnote{In principle we could have used OpenCell data to estimate the position of the towers in the RM dataset.
-However, the data does not provide the necessary information about the cell towers, such as network ID.}.
-We choose the MDC GPS logs as our source of location data for the cell towers for two reasons.
-First, as discussed above, there are more cells in the MDC dataset covered by the GPS logs than by the OpenCell database.
-Second, because the GPS logs measure the \textit{user's} location when he or she was connected to that tower, it provides a semantically more accurate value -- after all, as discussed in Section \ref{sec:intro}, the ultimate goal of the clustering algorithm is to provide data that will be fed to an algorithm that predicts \textit{user's} movements.
+In all experiments we used the same parameters value as the original Mobility 
+Profiler paper, namely $\delta_{duration} = 10 \text{minutes}$ and $\delta_{
+osc} = 3$.
 
@@
 
-\xxx{Compare MDC and Reality Mining, i.e. show that they are similar using the sampled RM?}
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{figs/cluster_size}
+\caption{The histogram of the size of clusters in the Reality Mining and 
+Nokia Mobile Data Challenge datasets. Because the MDC dataset originates from 
+sampling the user's phone every minute, we also sample the RM data to 
+visualize the impact of sampling. We see that both datasets produce clusters 
+that follow a similar size distribution.}
+\label{fig:cluster_size}
+\end{figure}
 
-We compare the Reality Mining dataset, which was originally used in the Mobility Profiler paper, with the Mobility Data Challenge.
-Table \xxx{Table with number of cells etc.} shows the number of cells, area codes and countries that are represented in each dataset.
-We can see that the MDC data has a much larger breadth in terms of the countries that are covered, but contains less cells and cell transitions.
-\xxx{Try to explain this difference.
-Talk about number of clusters.}
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{figs/cluster_quality}
+\caption{The histogram of the value of the quality metric for clusters in the 
+Reality Mining and Nokia Mobile Data Challenge datasets. Clusters that have a 
+quality metric of infinity are shown here as 0. We see that both datasets 
+produce clusters that follow a similar size distribution.}
+\label{fig:cluster_quality}
+\end{figure}
 
-However, both datasets show similar patterns in cluster size distribution, as evidenced in Figure \xxx{Figure with the oscillation histogram}, which also presents the cluster size distribution for the Reality Mining dataset sampled every minute.
-This is necessary for a fair comparison since the RM contains information on every handover, while the MDC samples the tower the phone is currently connected to every minute (so there is loss of information).
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{figs/cluster_area_density}
+\caption{The cummulative distribution function of the density of clusters in 
+an area code. We define the density as the number of clusters over the number 
+of cells in an area. An area with few, large clusters would have a density 
+close to zero, while an area with very few clusters will have a density close 
+to 1.}
+\label{fig:cluster_area_density}
+\end{figure}
 
-The fact that both datasets show a similar distribution is important for our results, since it allows us to extrapolate the results we find with the MDC dataset to the RM dataset with reasonable confidence.
+\begin{table}
+    \centering
+    \begin{tabular}{ | m{2.8cm} || r | m{3cm} | }
+    \hline
+    Criteria & Cells & Fraction of all cells in MDC dataset \\
+    \hline
+		With GPS data & 7508 & 31.75\%  \\
+    \hline
+		More than one GPS sample & 5622 & 23.78\% \\
+    \hline
+		More than five  GPS samples & 3644 & 15.41\% \\
+    \hline
+		GPS samples with  std. deviation~$ < 500$\,m & 3024 & 12.79\% \\
+    \hline
+		GPS samples with  std. deviation~$ < 150$\,m & 1173 & 4.96\% \\
+    \hline
+    \end{tabular}
+\caption{Statistics about the GPS data from the MDC GPS logs, for different 
+levels of data ``reliability''. In most of the paper we only consider the 
+location of a cell known when the GPS samples of that cell have standard 
+deviation less than 500\,m. }
+\label{tbl:gsm_logs_stats}
+\end{table}
 
-\subsection{Cluster statistics}
+We compare the Reality Mining dataset, which was originally used in the 
+Mobility Profiler paper, with the Mobility Data Challenge.
+Table~\ref{tbl:datasetComparison} shows the number of cells, area codes and countries 
+that are represented in each dataset. We can see that the MDC data has a much 
+larger breadth in terms of the country IDs that are covered\footnote{We do 
+not know the reason for this, since the researchers claim the data was 
+collected entirely in Lausanne. It is possible that the volunteers simply 
+travelled to many countries.}, but contains less cells and cell transitions.
 
-\xxx{Talk about cluster size, quality metric for both datasets}
+However, both datasets show similar patterns in cluster size distribution, as 
+evidenced in Figure~\ref{fig:cluster_size}, which also presents the cluster 
+size distribution for the Reality Mining dataset sampled every minute. This 
+is necessary for a fair comparison since the RM contains information on every 
+handover, while the MDC samples the tower the phone is currently connected to 
+every minute (so there is loss of information).
+
+We have also evaluated the cluster quality metric in both datasets, and 
+present the results in Figure~\ref{fig:cluster_quality}. Clusters whose cells 
+have no outgoing edges in the original oscillation graph have a quality 
+metric of infinity; these are represented as 0 in the histogram\footnote{This 
+is an unfortunate consequence of the quality metric adopted by the Mobility 
+Profiler paper. In section~\ref{sec:conclusion} we suggest an alternate 
+metric that is always between 0 and 1.}. We see that this is the case for the 
+vast majority of cells in both datasets.
+
+Finally, we also compare the density of clusters with respect to area codes 
+in Figure~\ref{fig:cluster_area_density}. We can also see that all datasets 
+follow similar patterns.
+
+The fact that both datasets show a similar distribution is important for our 
+results, since it allows us to extrapolate the results we find with the MDC 
+dataset to the RM dataset with reasonable confidence.
 
 \subsection{Location statistics}
+\label{sec:location_stats}
+Table~\ref{tbl:gsm_logs_stats} shows some statistics about the GPS 
+information obtained from the GPS logs in the MDC dataset. We use two 
+criteria to judge the confidence we have in a particular cell's position 
+estimate: the number of GPS samples for that cell and their standard 
+deviation. In the clustering efficiency analysis of
+Section~\ref{sec:effectiveness} we only consider cells that have more than one GPS sample 
+and whose samples' standard deviation is less than 500\,m. We choose these 
+parameters as they offer a good tradeoff between precision and sample size (
+for instance, choosing a stricter criteria of $\sigma < 150$ would leave us 
+with less than 5\% of all cells).
 
-\xxx{Talk about how many cells we have location data for, the difference between GPS logs and OpenCell, how many cells with $gps_{sightings} > 1$ and $stdddev < x$ etc.}
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{figs/mdc_open_distance}
+\caption{The histogram of the distance between cell tower positions estimated 
+using the GPS logs from the MDC dataset and the position in the open datasets.}
+\label{fig:mdc_open_distance}
+\end{figure}
 
+The data from the open dataset covers 5859 cells in the MDC dataset, or 24.78
+\%. We cannot use the data on the Reality Mining dataset, because the network 
+ID is not specified in its logs. Figure~\ref{fig:mdc_open_distance} shows a 
+histogram of the distance between the position of cell towers as computed by 
+MDC GPS logs and the open dataset. We see that the distance is usually small, 
+which reinfornces our confidence in the MDC GPS logs. We analyze how many 
+cells are covered by the combination of the open dataset and the GPS logs 
+from the MDC in Table~\ref{tbl:open_mdc_stats}. It is clear that a strikingly 
+small number of cells are present in both the MDC GPS logs and the open 
+dataset -- and the fraction becomes almost insignifcant when we restrict the 
+analysis to reliable GPS positions (those with small standard deviation in 
+the samples).
+\begin{table}
+    \centering
+    \begin{tabular}{ | m{2.8cm} || r | m{3cm} | }
+    \hline
+    Criteria & Cells & Fraction of all cells in MDC dataset \\
+    \hline
+		With GPS data from MDC and open datasets & 1645 & 6.96\%  \\
+    \hline
+		More than one GPS sample in MDC & 1220 & 5.16\% \\
+    \hline
+		More than five  GPS samples in MDC & 680 & 2.88\% \\
+    \hline
+		GPS samples in MDC with  std. deviation~$ < 500$\,m & 593 & 2.51\% \\
+    \hline
+		GPS samples in MDC with  std. deviation~$ < 150$\,m & 216 & 0.91\% \\
+    \hline
+    \end{tabular}
+\caption{Statistics for cells for which we have GPS data from the MDC GPS 
+logs \textit{and} the open datasets. We see that very few cells have a 
+reliable set of GPS samples in the MDC database and are present in the open 
+dataset. }
+\label{tbl:open_mdc_stats}
+\end{table}
+
+\begin{figure*}
+        \centering
+        \begin{subfigure}[t]{0.48\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{figs/cluster_distance_hist}
+                \caption{The histogram of the cluster distance for the MDC 
+dataset, using GPS data with $\sigma < 500$ from the MDC dataset. We only 
+consider clusters for which we have GPS for more than 70\% of the cells. We 
+observe that few clusters have a significant cluster distance.}
+                \label{fig:cluster_distance_hist}
+        \end{subfigure}%
+        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. 
+          %(or a blank line to force the subfigure onto a new line)
+        \begin{subfigure}[t]{0.48\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{figs/cluster_distance_cdf}
+                \caption{The CDF of the cluster distance under the same 
+conditions. Nearly all clusters have a distance metric less than 500\,m.}
+                \label{fig:cluster_distance_cdf}
+        \end{subfigure}
+				\caption{Cluster distance measurements.}\label{fig:cluster_distance}
+\end{figure*}
+
+\begin{figure*}
+        \centering
+        \begin{subfigure}[t]{0.48\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{figs/cluster_incomp_hist}
+                \caption{The histogram of the cluster incompleteness for the 
+MDC dataset, using GPS data with $\sigma < 500$ from the MDC dataset. A large 
+number of cells have an incompleteness measure of 0, which means all nearby 
+cells were clustered (or there weren't any nearby cells).}
+                \label{fig:cluster_incomp_hist}
+        \end{subfigure}%
+        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. 
+          %(or a blank line to force the subfigure onto a new line)
+        \begin{subfigure}[t]{0.48\textwidth}
+                \centering
+                \includegraphics[width=\textwidth]{figs/cluster_incomp_cdf}
+                \caption{The CDF of the cluster incompleteness under the same 
+conditions. Approximately 90\% of the cells have less than 5 nearby 
+unclustered neighbors.}
+                \label{fig:cluster_incomp_cdf}
+        \end{subfigure}
+				\caption{Cluster incompleteness measurements.}
+				\label{fig:cluster_incomp}
+\end{figure*}
 \subsection{Effectiveness of clustering algorithm}
+\label{sec:effectiveness}
+In this section we present the most important results of our project: an 
+empirical validation of the clustering algorithm from the point of view of 
+the efficiency of the clusters that it outputs.
 
-In this section we present the most important results of our project: an empirical validation of the clustering algorithm from the point of view of the efficiency of the clusters that it outputs.
-
+We are only able to experimentally evaluate the quality of clusters in the 
+MDC dataset, since it is the only one for which we have GPS information. We 
+choose the MDC GPS logs as our source of location data for the cell towers 
+for two reasons. First, as discussed in Section~\ref{sec:location_stats}, 
+there are more cells in the MDC dataset covered by the GPS logs than by the 
+open dataset. Second, because the GPS logs measure the \textit{user's} 
+location when he or she was connected to that tower, it provides a 
+semantically more accurate value -- after all, as discussed in
+Section~\ref{sec:intro}, the ultimate goal of the clustering algorithm is to provide data 
+that will be fed to an algorithm that predicts \textit{user's} movements.
 \subsubsection{Metrics}
 
-We define two metrics, cluster distance and cluster incompleteness.
+We define two metrics for assessing a cluster's quality: cluster distance and 
+cluster incompleteness.
 
 \paragraph{Cluster Distance} the average distance between cells in a cluster. 
-A good algorithm will group together towers that are geographically close and therefore clusters will have a low distance.
-The cluster distance is measured in meters.
+A good algorithm will group together towers that are geographically close and 
+therefore clusters will have a low distance. The cluster distance is measured 
+in meters.
 
-\paragraph{Cluster incompleteness} For a cell $k$, this is the number of cells within $\delta_{incomp}$ meters of it that are not in the same cluster as $k$.
-The algorithm is shown in \xxx{Show the cluster incompleteness algorithm.} Intuitively this measures if there are potentially oscillating towers that are not being clustered together.
-We can think of the cluster incompleteness as the dual of the cluster distance: the former favors aggressive algorithms while the latter encourages conservative approaches.
-Another way to see the benefit of using two metrics is noting that there are trivial solutions that minimize distance or incompleteness (each cell in its own cluster and one huge cluster with all cels, respectively), but there are no trivial solutions that minimize both metrics at the same time.
-
-\xxx{Talk about\begin{itemize}
-	\item Cluster distance histogram/CDF, average.
-	\item Mean cluster distance/cluster size (no correlation).
-	\item Cluster incompleteness CDF, average.
-	\item What can we conclude from this? That the algorithm works, pretty much.
-\end{itemize}}
+\paragraph{Cluster incompleteness} For a cell $k$, this is the number of 
+cells within $\delta_{incomp}$ meters of it that are not in the same cluster 
+as $k$. Intuitively this measures if there are potentially oscillating towers 
+that are not being clustered together. We can think of the cluster 
+incompleteness as the dual of the cluster distance: the former favors 
+aggressive algorithms while the latter encourages conservative approaches. 
+Another way to see the benefit of using two metrics is noting that there are 
+trivial solutions that minimize distance or incompleteness (each cell in its 
+own cluster and one huge cluster with all cels, respectively), but there are 
+no trivial solutions that minimize both metrics at the same time. We use
+$delta_{incomp} = 500$~meters.
 
@@
 
-In order to provide some confidence in the precision of the results, we only consider cells that were seen more than once in the GPS logs, with a standard deviation less than 500 meters.
-Only clusters in which more tha 70\% of cells meet this criteria are analyzed.
-This meant that only 82 clusters, or 9.44\% of the total, could be analyzed.
-Although this is a small fraction, there is no reason to think it is a biased sample if we assume there is no correlation between GPS precision and clustering inefficiency.
-Therefore, the results should hold for the entirety of the clusters, and, we conjecture, for other datasets too.
+In order to provide some confidence in the precision of the results, we only 
+consider cells that were seen more than once in the GPS logs, with a standard 
+deviation less than 500 meters. Only clusters in which more tha 70\% of cells 
+meet this criteria are analyzed. This meant that only 80 clusters, or 9.21\% 
+of the total, could be analyzed. Although this is a small fraction, there is 
+no reason to think it is a biased sample if we assume there is no correlation 
+between GPS precision and clustering inefficiency. Therefore, the results 
+should hold for the entirety of the clusters, and, we conjecture, for other 
+datasets too.
 
-Figures \xxx{cluster distance histogram} and \xxx{cluster distance CDF} show the histogram and CDF of the cluster distances. \xxx{Talk about that.}
-
-We also investigated whether there is a correlation between the size of a cluster and the its distance.
-A graph of both values is presented in Figure \xxx{size x distance}.
-We find that there is no significant correlation.
+Figure~\ref{fig:cluster_distance} show the histogram and CDF of the cluster 
+distances. The results confirm that the algorithm is clustering towers that 
+are largely close to each other: only one cluster (1.25\%) had a cluster 
+distance greater than 500\,m.
 
@@
 
-The measurements for cluster incompleteness are presented in Figures \xxx{cluster incompleteness histogram} and \xxx{cluster incompleteness CDF}.
-The average incompleteness for all cells with GPS data is 0.52 (standard deviation 1.37).
-This is a surprisingly low value and indicates that the algorithm works well.
+The measurements for cluster incompleteness are presented in
+Figure~\ref{fig:cluster_incomp}. The average incompleteness for all cells with GPS data 
+is 0.52 (standard deviation 1.37). This is a surprisingly low value and 
+indicates that the algorithm works well -- very few cells are not in the same 
+cluster as their nearby neighbors.
 
@@
 \label{sec:conclusion}
-We have evaluated a clustering algorithm used to eliminate the oscillation effect in cellular mobility prediction.
-Our main contribution is the use of GPS data to evaluate if the clustered towers are geographically close to each other, which would be circumstantial evidence in favor of the clustering algorithm.
-We find that the clustering algorithm proposed in \cite{mobilityprofiler} performs well on the two metrics we define, cluster distance and cluster incompleteness.
+We have evaluated a clustering algorithm used to eliminate the oscillation 
+effect in cellular mobility prediction. Our main contribution is the use of 
+GPS data to evaluate if the clustered towers are geographically close to each 
+other, which would be circumstantial evidence in favor of the clustering 
+algorithm. We find that the clustering algorithm proposed in \cite{
+mobilityprofiler} performs well on the two metrics we define, cluster 
+distance and cluster incompleteness.
 
-As part of future work, we would like to repeat the experiments here with more data.
-As discussed before, only 9.44\% of the clusters in the MDC dataset had enough information to be reliably processed by the algorithm.
-Another avenue of research is to investigate clustering algorithms that use a different clustering quality metric.
-One possible approach would be the modularity metric defined by Newman and Girvan~\cite{newman}.
-There is also a large body of literature on clustering in theoretical computer science, especially in social networks: Andersen and Peres find sparse cuts efficiently in~\cite{sparsecuts}, while Aroroa et.
-al try to find overlapping communities \cite{overlapping}.
-Finding overlapping communities has interesting applications for dealing with the oscillation effect, since towers can conceivably in more than one cluster.
+As part of future work, we would like to repeat the experiments here with 
+more data. As discussed before, only 9.44\% of the clusters in the MDC 
+dataset had enough information to be reliably processed by the algorithm. One 
+possible alternative is the professional cell tower location database 
+provided by Navizon~\cite{navizon}. Another avenue of research is to 
+investigate clustering algorithms that use a different clustering quality 
+metric. One possible approach would be the modularity metric defined by 
+Newman and Girvan~\cite{newman}. There is also a large body of literature on 
+clustering in theoretical computer science, especially in social networks: 
+Andersen and Peres find sparse cuts efficiently in~\cite{sparsecuts}, while 
+Aroroa et. al try to find overlapping communities \cite{overlapping}. Finding 
+overlapping communities has interesting applications for dealing with the 
+oscillation effect, since towers can conceivably in more than one cluster.
 

### 2012-12-13 docs/paper589.tex

@@
 literature that tries to counter this effect.  This algorithm tries to 
-cluster cell towers that users often oscillate between. In order to do 
+cluster cell towers that users often oscillate between. In order to 
 evaluate the algorithm's efficiency, we use a new dataset that allows us to 
-infer GPS data for cell towers. We find that the algorithm performs well.
+infer GPS data for cell towers. We find that the clustering algorithm performs well,
+as measured by two metrics we define.
 \end{abstract}
-\xxx{Make sure I didnt write Mobility Data Challenge (it should be Mobile 
-data challenge), or OpenCell (it should be open dataset).}
 
@@
 if they can forecast with reasonable accuracy
-where a user will be in the future \cite{LeapGraph} \cite{Liu97anoptimal}. By 
+where a user will be in the future \cite{LeapGraph}\cite{Liu97anoptimal}. By 
 being able to pre-allocate resources to towers that are most likely to
@@
 increases: indeed, in 2010 the amount of data traffic surpassed voice traffic 
-for the first time, and it is expected to double annually in the near future
+for the first time and it is expected to double annually in the near future
 \cite{EricssonData}.
 
-Another use of mobility prediction is to discover mobility profiles, i.e. 
-patterns in a user's movement \cite{mobilityprofiler}. As illustrated in \cite
+Another use of mobility prediction is to discover patterns in a user's movement, which is called a mobility profile\cite{mobilityprofiler}.
+As illustrated in \cite
 {mobilityprofiler}, mobility profiles have a wide range of possible 
@@
 
-The heart of every mobility prediction system based on cell tower connections 
+The main goal of every mobility prediction system based on cell tower connections 
 is to infer movement based on which cell the user's phone is connected to. 
@@
 towers. Figure~\ref{fig:stationary} shows an example of a stationary phone 
-that connects to different cell towers. This is a natural consequence of the 
+that connects to 5 different cell towers over the span of 20 minutes. This is a natural consequence of the 
 radio medium and the fact that towers very often have overlapping ranges. In 
@@
 The accuracy of current mobility prediction systems based on cell tower 
-connections is usually measured through machine learning techniques like cross
--validation that try to validate the ultimate goal of forecasting the user's 
+connections is usually measured through machine learning techniques such as cross-validation that try to validate the ultimate goal of forecasting the user's 
 future connections \cite{LeapGraph}. As far as we know there are no studies 
@@
 that are clustered together are actually close in real life. To evaluate 
-this, we use the Nokia Mobility Challenge dataset, which has data on both 
+this, our primary tool will be the Nokia Mobile Data Challenge dataset, which has data on both 
 cell tower connection (``GSM logs'') and the GPS position of the phones.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{figs/box_diagram}
+\caption{A high-level view of the components we describe in this report.
+Mobility paths are extracted from GSM connection logs.
+We then find pairs of cell towers that often oscillate between themselves and form a graph that represents this.
+We cluster this graph and consider every cluster to represent a group of oscillating towers.
+Finally, we use GPS data to evaluate the quality of those clusters.}
+\label{fig:box_diagram}
+\end{figure}
+
 The Mobility Profiler paper \cite{mobilityprofiler} describes an algorithm to 
 detect the oscillation effect. Our approach to evaluating this algorithm's 
-efficiency was the following. We implement the algorithm as described in \cite
+efficiency is the following. We implement the algorithm as described in \cite
 {mobilityprofiler} and test it in the same dataset used in the original 
@@
 users when they were connected to the towers. We therefore apply the 
-algorithm to another dataset, from the Nokia Mobility Challenge, in which the 
+algorithm to another dataset, from the Nokia Mobile Challenge, in which the 
 GPS position of the users is listed. In order to evaluate how successful the 
 clustering algorithm is, we devise two metrics: cluster distance and cluster 
-incompleteness.
+incompleteness. Figure~\ref{fig:box_diagram} shows a high-level view of our system.
 
@@
 set contains all towers that are within range of the phone's antenna at any 
-given time. Unfortunately, active set logs are not public which makes it 
+given time. Unfortunately, active set logs are not public, which makes it 
 impossible for us to evaluate the leap graph algorithm.
@@
 we use.}
+\label{fig:gsm_log_examples}
 \end{figure*}
@@
 transition events when a phone switches the tower it is connected to. These 
-events are timestamped, as shown in Figure~\ref{fig:gsm_log_example}. The 
+events are timestamped, as shown in Figure~\ref{fig:gsm_log_examples}. The 
 first step of the algorithm is to analyze the transitions and extract \textit{
@@
 at work, ignoring the oscillation effect temporarily, the phone will be 
-connected to a single tower. Finally, when the user the goes home the 
-handovers that occur during the drive will belong to a different mobility path.
+connected to a single tower. Finally, when the user goes home the 
+handovers that occur during that drive will belong to a different mobility path.
 
-More formally, let $\Delta_i$ be the time difference between two consecutive 
+Formally, let $\Delta_i$ be the time difference between two consecutive 
 handovers for a particular user's connection log. In other words, $\Delta_i$ 
@@
 that switches do not need to be consecutive. The authors argue that this 
-allows for detection of overlapping towers in dense region where multiple 
+allows for detection of overlapping towers in dense regions where multiple 
 towers may be within range at the same time.
@@
 step is to build the oscillation graph. The set $V$ of vertices in the graph 
-are all cell towers in the dataset, and the weighted edges in $E$ connect 
-oscillating pairs. Let $P_i$ be a mobility path.The weight of every edge, 
+are all cell towers in the dataset, and weighted edges in $E$ connect 
+oscillating pairs. Let $P_i$ be a mobility path. The weight of every edge, 
 which we also call its \textit{support}, is defined as:
@@
 \begin{equation*}
-s(x,y) = \frac{|\left\{P_i| (x, y) \in P_i \wedge (x, y) \text{oscillated}
+s(x,y) = \frac{|\left\{P_i| (x, y) \in P_i \wedge (x, y) \text{ oscillated}
 \right\}|}{|\left\{P_i| (x, y) \in P_i\right\}|}
@@
 In words, $s(x,y)$ is the ratio of how many times the cell towers $x$ and $y$ 
-oscillated over the number of mobility paths in which $x$ and $y$ appear. For 
-example, a pair of towers such that in \textit{every} mobility path that both 
-of them appear the switch count is greater than $k$ would have a support 
+oscillated over the number of mobility paths in which $x$ and $y$ appear together. For 
+example, a pair of towers such that their switch count is greater than $k$ in \textit{every} mobility path that both of appear in would have a support 
 equal to 1. If a pair only switches more or equal to $k$ times in half the 
@@
 Intuitively, a cluster represents a group of towers that are not indicative 
-of actual user movement. The Mobility Profiler paper uses an greedy divise 
+of actual user movement. The Mobility Profiler paper uses a greedy divise 
 algorithm for clustering. At every iteration, the algorithm removes the edge 
@@
 graph. Formally, let $C$ be a cluster with the set of edges $E$, and $E_{out}$
- be the set of edges that have exactly one vertex in $C$, and define $w(e)$ 
-to be an edge's weight:
+ be the set of edges in the original graph that have exactly one vertex in $C$, and call an edge's weight $w(e)$. Then the quality of $C$ is:
 
@@
 the graph. The algorithm continues until there are no more vertices in the 
-graph. The algorithm outputs all the clusters it found. Our main goal is to 
+graph. The algorithm outputs all the clusters it finds. Our main goal is to 
 evaluate these clusters according to the GPS position of the users when they 
@@
 hours of data logged from 94 volunteers' cell phones during 2004. The breadth 
-of information collected is impressive, ranging from cell tower IDs to which 
+of information collected ranges from cell tower IDs to which 
 Bluetooth devices are within 5 meters of the users. This dataset has been 
 widely used in research into social and friendship networks. For our project, 
-following~\cite{mobilityprofiler}, the focus is the cell tower connection logs.
+following~\cite{mobilityprofiler}, the focus is in cell tower connection logs.
 
@@
 Our main contribution is an analysis of the oscillation clustering algorithm 
-using GPS data. We use two sources of GPS data, the Nokia MDC and the 
-OpenCell database.
+using GPS data. We use two sources of GPS data, the Nokia MDC and the combination of two open datasets.
 
-\subsubsection{Nokia Mobility Data Challenge -- GPS}
+\subsubsection{Nokia Mobile Data Challenge -- GPS}
 
@@
 
-We implemented the clustering algorithm in \cite{mobilityprofiler} for the 
-Reality Mining and the Nokia Mobility Data Challenge datasets and compared 
+We implement the clustering algorithm in \cite{mobilityprofiler} for the 
+Reality Mining and the Nokia Mobile Data Challenge datasets\footnote{Our implementation is written in 5500 lines of Java and uses the JGraphT, Hibernate and H2 libraries.
+It is freely available at https://github.com/pdaquino/EECS589.} and compare
 the results of the clustering in both datasets (adjusting for the fact that 
-the MDC dataset's data is sampled every minute). In this section, we report 
-the results of running the algorithm in those datasets and analyze the 
-results. We find that the Reality Mining and MDC datasets have similar 
-patterns of cluster size and density. We also used GPS data to evaluate the 
+the MDC dataset's data is sampled every minute). We find that the Reality Mining and MDC datasets have similar 
+patterns of cluster size and density. We also use GPS data to evaluate the 
 accuracy and quality of the resulting clusters. Our experiments show that the 
@@
 We compare the Reality Mining dataset, which was originally used in the 
-Mobility Profiler paper, with the Mobility Data Challenge.
+Mobility Profiler paper, with the Mobile Data Challenge.
 Table~\ref{tbl:datasetComparison} shows the number of cells, area codes and countries 
@@
 
-We have also evaluated the cluster quality metric in both datasets, and 
+We also evaluate the cluster quality metric in both datasets, and 
 present the results in Figure~\ref{fig:cluster_quality}. Clusters whose cells 
@@
 
-\subsection{Location statistics}
+\subsection{Cell location statistics}
 \label{sec:location_stats}
-Table~\ref{tbl:gsm_logs_stats} shows some statistics about the GPS 
+
+Table~\ref{tbl:gsm_logs_stats} shows statistics about the GPS 
 information obtained from the GPS logs in the MDC dataset. We use two 
@@
 Section~\ref{sec:effectiveness} we only consider cells that have more than one GPS sample 
-and whose samples' standard deviation is less than 500\,m. We choose these 
+and whose standard deviation is less than 500\,m. We choose these 
 parameters as they offer a good tradeoff between precision and sample size (
@@
 
-The data from the open dataset covers 5859 cells in the MDC dataset, or 24.78
+The data from the open dataset covers 5859 cells in the MDC GSM dataset, or 24.78
 \%. We cannot use the data on the Reality Mining dataset, because the network 
@@
 
-\paragraph{Cluster Distance} the average distance between cells in a cluster. 
+\paragraph{Cluster Distance} The average distance between cells in a cluster. 
 A good algorithm will group together towers that are geographically close and 
@@
 cells within $\delta_{incomp}$ meters of it that are not in the same cluster 
-as $k$. Intuitively this measures if there are potentially oscillating towers 
+as $k$. Intuitively, this measures if there are potentially oscillating towers 
 that are not being clustered together. We can think of the cluster 
 incompleteness as the dual of the cluster distance: the former favors 
-aggressive algorithms while the latter encourages conservative approaches. 
+aggressive algorithms that overestimate the size of clusters while the latter encourages conservative approaches. 
 Another way to see the benefit of using two metrics is noting that there are 
@@
 no trivial solutions that minimize both metrics at the same time. We use
-$delta_{incomp} = 500$~meters.
+$\delta_{incomp} = 500$~meters.
 
-\subsubsection{Cluster distance}
+\subsubsection{Cluster distance results}
 
@@
 consider cells that were seen more than once in the GPS logs, with a standard 
-deviation less than 500 meters. Only clusters in which more tha 70\% of cells 
-meet this criteria are analyzed. This meant that only 80 clusters, or 9.21\% 
+deviation less than 500 meters. Only clusters in which more than 70\% of cells 
+meet this criteria are analyzed. This means that only 80 clusters, or 9.21\% 
 of the total, could be analyzed. Although this is a small fraction, there is 
@@
 
-\subsubsection{Cluster incompleteness}
+\subsubsection{Cluster incompleteness results}
 
@@
 As part of future work, we would like to repeat the experiments here with 
-more data. As discussed before, only 9.44\% of the clusters in the MDC 
+more data. As discussed before, only 9.21\% of the clusters in the MDC 
 dataset had enough information to be reliably processed by the algorithm. One 
